### 2015-11-03 PrincipiaMetaphysicaAutomationPart1/doc/document/TheoryOfAbstractObjectsV1.tex
some further tests

%
\begin{isabellebody}%
\def\isabellecontext{TheoryOfAbstractObjectsV{\isadigit{1}}}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isamarkupsection{Introduction%
}
\isamarkuptrue%
%
\begin{isamarkuptext}%
This work is related to \cite{J23}, which is significantly extends ...%
\end{isamarkuptext}%
\isamarkuptrue%
%
\isamarkupsection{Preliminaries%
}
\isamarkuptrue%
\isacommand{typedecl}\isamarkupfalse%
\ i\isanewline
%
\isamarkupcmt{the type possible worlds; the formalism explicitly encodes Kripke style semantics%
}
\isanewline
\isacommand{type{\isacharunderscore}synonym}\isamarkupfalse%
\ io\ {\isacharequal}\ {\isachardoublequoteopen}{\isacharparenleft}i\ {\isasymRightarrow}\ bool{\isacharparenright}{\isachardoublequoteclose}\ \isanewline
%
\isamarkupcmt{formulas are essentially of this type%
}
\isanewline
%
\isamarkupcmt{predicates on worlds%
}
\isanewline
\isacommand{typedecl}\isamarkupfalse%
\ e\isanewline
%
\isamarkupcmt{the raw type of entities/objects (abstract or ordinary)%
}
\isanewline
\isacommand{datatype}\isamarkupfalse%
\ {\isacharprime}a\ opt\ {\isacharequal}\ Error\ {\isacharprime}a\ {\isacharbar}\ Term\ {\isacharprime}a\ {\isacharbar}\ Form\ {\isacharprime}a\ {\isacharbar}\ PropForm\ {\isacharprime}a\isanewline
\isanewline
\isanewline
\isacommand{consts}\isamarkupfalse%
\ cw\ {\isacharcolon}{\isacharcolon}\ i\ \isanewline
%
\isamarkupcmt{the distinguished actual world%
}
\isanewline
\isacommand{consts}\isamarkupfalse%
\ dE{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}e{\isachardoublequoteclose}\ dIO{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}io{\isachardoublequoteclose}\ dEIO{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}e{\isasymRightarrow}io{\isachardoublequoteclose}\ dEEIO{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}e{\isasymRightarrow}e{\isasymRightarrow}io{\isachardoublequoteclose}\ dEEEIO{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}e{\isasymRightarrow}e{\isasymRightarrow}e{\isasymRightarrow}io{\isachardoublequoteclose}\ dA{\isacharcolon}{\isacharcolon}{\isacharprime}a\isanewline
%
\isamarkupcmt{some fixed dummy symbols; we anyway assume that the domains are non-empty%
}
\isanewline
%
\isamarkupcmt{needed as dummy object in some cases below%
}
%
\begin{isamarkuptext}%
Meta-logical predicates.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ isWff\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharprime}a\ opt{\isasymRightarrow}bool{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}isWff\ {\isasymphi}\ {\isasymequiv}\ case\ {\isasymphi}\ of\ Error\ {\isasympsi}\ {\isasymRightarrow}\ False\ {\isacharbar}\ Term\ {\isasympsi}\ {\isasymRightarrow}\ False\ {\isacharbar}{\isacharunderscore}\ {\isasymRightarrow}\ True{\isachardoublequoteclose}\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ isForm\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharprime}a\ opt{\isasymRightarrow}bool{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}isForm\ {\isasymphi}\ {\isasymequiv}\ case\ {\isasymphi}\ of\ Form\ {\isasympsi}\ {\isasymRightarrow}\ True\ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ False{\isachardoublequoteclose}\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ isPropForm\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharprime}a\ opt{\isasymRightarrow}bool{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}isPropForm\ {\isasymphi}\ {\isasymequiv}\ case\ {\isasymphi}\ of\ PropForm\ {\isasympsi}\ {\isasymRightarrow}\ True\ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ False{\isachardoublequoteclose}\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ isTerm\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharprime}a\ opt{\isasymRightarrow}bool{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}isTerm\ {\isasymphi}\ {\isasymequiv}\ case\ {\isasymphi}\ of\ Term\ {\isasympsi}\ {\isasymRightarrow}\ True\ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ False{\isachardoublequoteclose}\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ isError\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}{\isacharprime}a\ opt{\isasymRightarrow}bool{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}isError\ {\isasymphi}\ {\isasymequiv}\ case\ {\isasymphi}\ of\ Error\ {\isasympsi}\ {\isasymRightarrow}\ True\ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ False{\isachardoublequoteclose}\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ valid\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}io\ opt{\isasymRightarrow}bool{\isachardoublequoteclose}\ \ \isakeyword{where}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymphi}{\isacharbrackright}\ {\isasymequiv}\ case\ {\isasymphi}\ of\ \isanewline
\ \ \ \ PropForm\ {\isasympsi}\ {\isasymRightarrow}\ {\isasymforall}w{\isachardot}{\isacharparenleft}{\isasympsi}\ w{\isacharparenright}\isanewline
\ \ {\isacharbar}\ Form\ {\isasympsi}\ {\isasymRightarrow}\ {\isasymforall}w{\isachardot}{\isacharparenleft}{\isasympsi}\ w{\isacharparenright}\isanewline
\ \ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ False{\isachardoublequoteclose}\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ satisfiable\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}io\ opt{\isasymRightarrow}bool{\isachardoublequoteclose}\ \ \isakeyword{where}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymphi}{\isacharbrackright}\isactrlsup s\isactrlsup a\isactrlsup t\ {\isasymequiv}\ case\ {\isasymphi}\ of\ \isanewline
\ \ \ \ PropForm\ {\isasympsi}\ {\isasymRightarrow}\ {\isasymexists}w{\isachardot}{\isacharparenleft}{\isasympsi}\ w{\isacharparenright}\isanewline
\ \ {\isacharbar}\ Form\ {\isasympsi}\ {\isasymRightarrow}\ {\isasymexists}w{\isachardot}{\isacharparenleft}{\isasympsi}\ w{\isacharparenright}\isanewline
\ \ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ False{\isachardoublequoteclose}\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ countersatifiable\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}io\ opt{\isasymRightarrow}bool{\isachardoublequoteclose}\ \ \isakeyword{where}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymphi}{\isacharbrackright}\isactrlsup c\isactrlsup s\isactrlsup a\isactrlsup t\ {\isasymequiv}\ \ case\ {\isasymphi}\ of\ \isanewline
\ \ \ \ PropForm\ {\isasympsi}\ {\isasymRightarrow}\ {\isasymexists}w{\isachardot}{\isasymnot}{\isacharparenleft}{\isasympsi}\ w{\isacharparenright}\isanewline
\ \ {\isacharbar}\ Form\ {\isasympsi}\ {\isasymRightarrow}\ {\isasymexists}w{\isachardot}{\isasymnot}{\isacharparenleft}{\isasympsi}\ w{\isacharparenright}\isanewline
\ \ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ False{\isachardoublequoteclose}\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ invalid\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}io\ opt{\isasymRightarrow}bool{\isachardoublequoteclose}\ \ \isakeyword{where}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymphi}{\isacharbrackright}\isactrlsup i\isactrlsup n\isactrlsup v\ {\isasymequiv}\ case\ {\isasymphi}\ of\ \isanewline
\ \ \ \ PropForm\ {\isasympsi}\ {\isasymRightarrow}\ {\isasymforall}w{\isachardot}{\isasymnot}{\isacharparenleft}{\isasympsi}\ w{\isacharparenright}\isanewline
\ \ {\isacharbar}\ Form\ {\isasympsi}\ {\isasymRightarrow}\ {\isasymforall}w{\isachardot}{\isasymnot}{\isacharparenleft}{\isasympsi}\ w{\isacharparenright}\isanewline
\ \ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ False{\isachardoublequoteclose}%
\isamarkupsection{Encoding of the language%
}
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ {\isasymA}{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}io\ opt\ {\isasymRightarrow}\ io\ opt{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymA}\ {\isasymphi}\ {\isasymequiv}\ case\ {\isasymphi}\ of\ \isanewline
\ \ \ \ Form\ {\isasympsi}\ {\isasymRightarrow}\ Form\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasympsi}\ cw{\isacharparenright}\isanewline
\ \ {\isacharbar}\ PropForm\ {\isasympsi}\ {\isasymRightarrow}\ PropForm\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasympsi}\ cw{\isacharparenright}\isanewline
\ \ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ Error\ dIO{\isachardoublequoteclose}%
\begin{isamarkuptext}%
actuality operator; \isa{{\isasymphi}} is evaluated wrt the current world; Error is passed on%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ Enc{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}e\ opt{\isasymRightarrow}{\isacharparenleft}e{\isasymRightarrow}io{\isacharparenright}\ opt{\isasymRightarrow}io\ opt{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isacharless}x{\isasymcirc}P{\isachargreater}\ {\isasymequiv}\ case\ {\isacharparenleft}x{\isacharcomma}P{\isacharparenright}\ of\ \isanewline
\ \ \ \ {\isacharparenleft}Term\ y{\isacharcomma}Term\ Q{\isacharparenright}\ {\isasymRightarrow}\ Form\ {\isacharparenleft}{\isasymlambda}w{\isachardot}{\isacharparenleft}Q\ y{\isacharparenright}\ w{\isacharparenright}\ \ \ \isanewline
\ \ {\isacharbar}\ {\isacharparenleft}{\isacharunderscore}{\isacharcomma}{\isacharunderscore}{\isacharparenright}\ {\isasymRightarrow}\ Error\ dIO{\isachardoublequoteclose}%
\begin{isamarkuptext}%
\isa{{\isasymkappa}\isactrlsub {\isadigit{1}}{\isasymPi}\isactrlsup {\isadigit{1}}} will be written here as \isa{{\isacharless}{\isasymkappa}\isactrlsub {\isadigit{1}}{\isasymcirc}{\isasymPi}\isactrlsup {\isadigit{1}}{\isachargreater}}; \isa{{\isasymkappa}\isactrlsub {\isadigit{1}}{\isasymPi}\isactrlsup {\isadigit{1}}} is a Form; Error is passed on%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ Exe{\isadigit{1}}{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}e{\isasymRightarrow}io{\isacharparenright}\ opt{\isasymRightarrow}e\ opt{\isasymRightarrow}io\ opt{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isacharless}P{\isasymbullet}x{\isachargreater}\ {\isasymequiv}\ case\ {\isacharparenleft}P{\isacharcomma}x{\isacharparenright}\ of\ \isanewline
\ \ \ \ {\isacharparenleft}Term\ Q{\isacharcomma}Term\ y{\isacharparenright}\ {\isasymRightarrow}\ PropForm\ {\isacharparenleft}{\isasymlambda}w{\isachardot}{\isacharparenleft}Q\ y{\isacharparenright}\ w{\isacharparenright}\isanewline
\ \ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ Error\ dIO{\isachardoublequoteclose}%
\begin{isamarkuptext}%
\isa{{\isasymPi}\isactrlsup {\isadigit{1}}{\isasymkappa}\isactrlsub {\isadigit{1}}}  will be written here as \isa{{\isacharless}{\isasymPi}\isactrlsup {\isadigit{2}}{\isasymbullet}{\isasymkappa}\isactrlsub {\isadigit{1}}{\isachargreater}}; \isa{{\isasymPi}\isactrlsup {\isadigit{1}}{\isasymkappa}\isactrlsub {\isadigit{1}}} is a PropForm; Error is passed on%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ Exe{\isadigit{2}}{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}e{\isasymRightarrow}e{\isasymRightarrow}io{\isacharparenright}\ opt{\isasymRightarrow}e\ opt{\isasymRightarrow}e\ opt{\isasymRightarrow}io\ opt{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isacharless}P{\isasymbullet}x{\isadigit{1}}{\isacharcomma}x{\isadigit{2}}{\isachargreater}\ {\isasymequiv}\ case\ {\isacharparenleft}P{\isacharcomma}x{\isadigit{1}}{\isacharcomma}x{\isadigit{2}}{\isacharparenright}\ of\ \isanewline
\ \ \ \ {\isacharparenleft}Term\ Q{\isacharcomma}Term\ y{\isadigit{1}}{\isacharcomma}Term\ y{\isadigit{2}}{\isacharparenright}\ {\isasymRightarrow}\ PropForm\ {\isacharparenleft}{\isasymlambda}w{\isachardot}{\isacharparenleft}Q\ y{\isadigit{1}}\ y{\isadigit{2}}{\isacharparenright}\ w{\isacharparenright}\isanewline
\ \ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ Error\ dIO{\isachardoublequoteclose}%
\begin{isamarkuptext}%
\isa{{\isasymPi}\isactrlsup {\isadigit{2}}{\isasymkappa}\isactrlsub {\isadigit{1}}{\isasymkappa}\isactrlsub {\isadigit{2}}}   will be written here as \isa{{\isacharless}{\isasymPi}\isactrlsup {\isadigit{2}}{\isasymbullet}{\isasymkappa}\isactrlsub {\isadigit{1}}{\isacharcomma}{\isasymkappa}\isactrlsub {\isadigit{2}}{\isachargreater}}; \isa{{\isasymPi}\isactrlsup {\isadigit{2}}{\isasymkappa}\isactrlsub {\isadigit{1}}{\isasymkappa}\isactrlsub {\isadigit{2}}} is a PropForm; Error is passed on%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ Exe{\isadigit{3}}{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}e{\isasymRightarrow}e{\isasymRightarrow}e{\isasymRightarrow}io{\isacharparenright}\ opt{\isasymRightarrow}e\ opt{\isasymRightarrow}e\ opt{\isasymRightarrow}e\ opt{\isasymRightarrow}io\ opt{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isacharless}P{\isasymbullet}x{\isadigit{1}}{\isacharcomma}x{\isadigit{2}}{\isacharcomma}x{\isadigit{3}}{\isachargreater}\ {\isasymequiv}\ case\ {\isacharparenleft}P{\isacharcomma}x{\isadigit{1}}{\isacharcomma}x{\isadigit{2}}{\isacharcomma}x{\isadigit{3}}{\isacharparenright}\ of\ \isanewline
\ \ \ \ {\isacharparenleft}Term\ Q{\isacharcomma}Term\ y{\isadigit{1}}{\isacharcomma}Term\ y{\isadigit{2}}{\isacharcomma}Term\ y{\isadigit{3}}{\isacharparenright}\ {\isasymRightarrow}\ PropForm\ {\isacharparenleft}{\isasymlambda}w{\isachardot}{\isacharparenleft}Q\ y{\isadigit{1}}\ y{\isadigit{2}}\ y{\isadigit{3}}{\isacharparenright}\ w{\isacharparenright}\isanewline
\ \ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ Error\ dIO{\isachardoublequoteclose}%
\begin{isamarkuptext}%
\isa{{\isasymPi}\isactrlsup {\isadigit{3}}\isactrlsub {\isadigit{1}}{\isasymkappa}\isactrlsub {\isadigit{2}}{\isasymkappa}\isactrlsub {\isadigit{3}}}  will be written here as \isa{{\isacharless}{\isasymPi}\isactrlsup {\isadigit{2}}{\isasymbullet}{\isasymkappa}\isactrlsub {\isadigit{1}}{\isacharcomma}{\isasymkappa}\isactrlsub {\isadigit{2}}{\isacharcomma}{\isasymkappa}\isactrlsub {\isadigit{3}}{\isachargreater}}; \isa{{\isasymPi}\isactrlsup {\isadigit{3}}\isactrlsub {\isadigit{1}}{\isasymkappa}\isactrlsub {\isadigit{2}}{\isasymkappa}\isactrlsub {\isadigit{3}}} is a PropForm; Error is passed on; 
we could, of course, introduce further operators: Exe4, Exe5, etc.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ z{\isacharunderscore}not{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}io\ opt{\isasymRightarrow}io\ opt{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymnot}\isactrlsup z\ {\isasymphi}\ {\isasymequiv}\ case\ {\isasymphi}\ of\ \isanewline
\ \ \ \ Form\ {\isasympsi}\ {\isasymRightarrow}\ Form\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymnot}\ {\isasympsi}\ w{\isacharparenright}\isanewline
\ \ {\isacharbar}\ PropForm\ {\isasympsi}\ {\isasymRightarrow}\ PropForm\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymnot}\ {\isasympsi}\ w{\isacharparenright}\isanewline
\ \ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ Error\ dIO{\isachardoublequoteclose}%
\begin{isamarkuptext}%
negation operator; \isa{{\isasymnot}\isactrlsup z\ {\isasymphi}} inherits its type from \isa{{\isasymphi}} if \isa{{\isasymphi}} is Form or PropForm; 
Error is passed on%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ z{\isacharunderscore}implies{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}io\ opt{\isasymRightarrow}io\ opt{\isasymRightarrow}io\ opt{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymrightarrow}\isactrlsup z\ {\isasympsi}\ {\isasymequiv}\ case\ {\isacharparenleft}{\isasymphi}{\isacharcomma}{\isasympsi}{\isacharparenright}\ of\ \isanewline
\ \ \ \ {\isacharparenleft}PropForm\ {\isasymalpha}{\isacharcomma}PropForm\ {\isasymbeta}{\isacharparenright}\ {\isasymRightarrow}\ PropForm\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymalpha}\ w\ {\isasymlongrightarrow}\ {\isasymbeta}\ w{\isacharparenright}\isanewline
\ \ {\isacharbar}\ {\isacharparenleft}Form\ {\isasymalpha}{\isacharcomma}Form\ {\isasymbeta}{\isacharparenright}\ {\isasymRightarrow}\ Form\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymalpha}\ w\ {\isasymlongrightarrow}\ {\isasymbeta}\ w{\isacharparenright}\isanewline
\ \ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ Error\ dIO{\isachardoublequoteclose}%
\begin{isamarkuptext}%
implication operator; \isa{{\isasymphi}\ {\isasymrightarrow}\isactrlsup z\ {\isasympsi}} returns returns a PropForm if both are PropForms, Form if both are Forms,
otherwise it returns Error%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ z{\isacharunderscore}forall{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}{\isacharprime}a{\isasymRightarrow}io\ opt{\isacharparenright}{\isasymRightarrow}io\ opt{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymforall}\ {\isasymPhi}\ {\isasymequiv}\ case\ {\isacharparenleft}{\isasymPhi}\ dA{\isacharparenright}\ of\isanewline
\ \ \ \ PropForm\ {\isasymphi}\ {\isasymRightarrow}\ PropForm\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}x{\isachardot}\ case\ {\isacharparenleft}{\isasymPhi}\ x{\isacharparenright}\ of\ PropForm\ {\isasympsi}\ {\isasymRightarrow}\ {\isasympsi}\ w{\isacharparenright}\isanewline
\ \ {\isacharbar}\ Form\ {\isasymphi}\ {\isasymRightarrow}\ Form\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}x{\isachardot}\ case\ {\isacharparenleft}{\isasymPhi}\ x{\isacharparenright}\ of\ Form\ {\isasympsi}\ {\isasymRightarrow}\ {\isasympsi}\ w{\isacharparenright}\isanewline
\ \ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ Error\ dIO{\isachardoublequoteclose}%
\begin{isamarkuptext}%
universal quantification; \isa{{\isasymforall}{\isacharparenleft}{\isasymlambda}x{\isachardot}{\isasymphi}{\isacharparenright}} inherits its kind (Form or PropForm) from \isa{{\isasymphi}}; Error is passed on
\isa{{\isasymforall}{\isacharparenleft}{\isasymlambda}x{\isachardot}{\isasymphi}{\isacharparenright}} is mapped to \isa{{\isacharparenleft}{\isasymlambda}w{\isachardot}{\isasymforall}x{\isachardot}{\isasymphi}xw{\isacharparenright}} as expected%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ z{\isacharunderscore}box{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}io\ opt{\isasymRightarrow}io\ opt{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymbox}\ {\isasymphi}\ {\isasymequiv}\ case\ {\isasymphi}\ of\ \isanewline
\ \ \ \ Form\ {\isasympsi}\ {\isasymRightarrow}\ Form\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}v{\isachardot}\ {\isasympsi}\ v{\isacharparenright}\isanewline
\ \ {\isacharbar}\ PropForm\ {\isasympsi}\ {\isasymRightarrow}\ PropForm\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymforall}v{\isachardot}\ {\isasympsi}\ v{\isacharparenright}\isanewline
\ \ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ Error\ dIO{\isachardoublequoteclose}%
\begin{isamarkuptext}%
box operator; \isa{{\isasymbox}\ {\isasymphi}} inherits its type (Form or PropForm) from \isa{{\isasymphi}}; Error is passed on.
Note that the \isa{{\isasymbox}}-operator is defined here without an accessibility relation; this is ok since we assume logic S5.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ lam{\isadigit{0}}{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}io\ opt{\isasymRightarrow}io\ opt{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymlambda}\isactrlsup {\isadigit{0}}\ {\isasymphi}\ {\isasymequiv}\ case\ {\isasymphi}\ of\ \isanewline
\ \ \ \ PropForm\ {\isasympsi}\ {\isasymRightarrow}\ PropForm\ {\isasympsi}\isanewline
\ \ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ Error\ dIO{\isachardoublequoteclose}%
\begin{isamarkuptext}%
0-arity lambda abstraction; \isa{{\isasymlambda}\isactrlsup {\isadigit{0}}\ {\isasymphi}} returns PropForm \isa{{\isasymphi}} if \isa{{\isasymphi}} is a PropForm, otherwise Error%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ lam{\isadigit{1}}{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}e{\isasymRightarrow}io\ opt{\isacharparenright}{\isasymRightarrow}{\isacharparenleft}e{\isasymRightarrow}io{\isacharparenright}\ opt{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymlambda}\isactrlsup {\isadigit{1}}\ {\isasymPhi}\ {\isasymequiv}\ case\ {\isacharparenleft}{\isasymPhi}\ dE{\isacharparenright}\ of\isanewline
\ \ \ \ PropForm\ {\isasymphi}\ {\isasymRightarrow}\ Term\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ case\ {\isacharparenleft}{\isasymPhi}\ x{\isacharparenright}\ of\ PropForm\ {\isasymphi}\ {\isasymRightarrow}\ {\isasymphi}{\isacharparenright}\isanewline
\ \ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ Error\ {\isacharparenleft}{\isasymlambda}x{\isachardot}\ dIO{\isacharparenright}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
1-arity lambda abstraction; \isa{{\isasymlambda}\isactrlsup {\isadigit{1}}{\isacharparenleft}{\isasymlambda}x{\isachardot}{\isasymphi}{\isacharparenright}} returns Term \isa{{\isacharparenleft}{\isasymlambda}x{\isachardot}{\isasymphi}{\isacharparenright}} if \isa{{\isasymphi}} is a PropForm, 
otherwise Error%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ lam{\isadigit{2}}{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}e{\isasymRightarrow}e{\isasymRightarrow}io\ opt{\isacharparenright}{\isasymRightarrow}{\isacharparenleft}e{\isasymRightarrow}e{\isasymRightarrow}io{\isacharparenright}\ opt{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymlambda}\isactrlsup {\isadigit{2}}\ {\isasymPhi}\ {\isasymequiv}\ case\ {\isacharparenleft}{\isasymPhi}\ dE\ dE{\isacharparenright}\ of\isanewline
\ \ \ \ PropForm\ {\isasymphi}\ {\isasymRightarrow}\ Term\ {\isacharparenleft}{\isasymlambda}x\ y{\isachardot}\ case\ {\isacharparenleft}{\isasymPhi}\ x\ y{\isacharparenright}\ of\ PropForm\ {\isasymphi}\ {\isasymRightarrow}\ {\isasymphi}{\isacharparenright}\isanewline
\ \ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ Error\ {\isacharparenleft}{\isasymlambda}x\ y{\isachardot}\ dIO{\isacharparenright}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
2-arity lambda abstraction; \isa{{\isasymlambda}\isactrlsup {\isadigit{2}}{\isacharparenleft}{\isasymlambda}xy{\isachardot}{\isasymphi}{\isacharparenright}} returns Term \isa{{\isacharparenleft}{\isasymlambda}xy{\isachardot}{\isasymphi}{\isacharparenright}} if \isa{{\isasymphi}} is a PropForm, 
otherwise Error%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ lam{\isadigit{3}}{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}e{\isasymRightarrow}e{\isasymRightarrow}e{\isasymRightarrow}io\ opt{\isacharparenright}{\isasymRightarrow}{\isacharparenleft}e{\isasymRightarrow}e{\isasymRightarrow}e{\isasymRightarrow}io{\isacharparenright}\ opt{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymlambda}\isactrlsup {\isadigit{3}}\ {\isasymPhi}\ {\isasymequiv}\ case\ {\isacharparenleft}{\isasymPhi}\ dE\ dE\ dE{\isacharparenright}\ of\isanewline
\ \ \ \ PropForm\ {\isasymphi}\ {\isasymRightarrow}\ Term\ {\isacharparenleft}{\isasymlambda}x\ y\ z{\isachardot}\ case\ {\isacharparenleft}{\isasymPhi}\ x\ y\ z{\isacharparenright}\ of\ PropForm\ {\isasymphi}\ {\isasymRightarrow}\ {\isasymphi}{\isacharparenright}\isanewline
\ \ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ Error\ {\isacharparenleft}{\isasymlambda}x\ y\ z{\isachardot}\ dIO{\isacharparenright}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
3-arity lambda abstraction; \isa{{\isasymlambda}\isactrlsup {\isadigit{2}}{\isacharparenleft}{\isasymlambda}xyz{\isachardot}{\isasymphi}{\isacharparenright}} returns Term \isa{{\isacharparenleft}{\isasymlambda}xyz{\isachardot}{\isasymphi}{\isacharparenright}} if \isa{{\isasymphi}} is a PropForm, 
otherwise Error; we could, of course, introduce further operators: \isa{{\isasymlambda}\isactrlsup {\isadigit{4}}}, \isa{{\isasymlambda}\isactrlsup {\isadigit{5}}}, etc.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ that{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}e{\isasymRightarrow}io\ opt{\isacharparenright}{\isasymRightarrow}e\ opt{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymepsilon}\ {\isasymPhi}\ {\isasymequiv}\ case\ {\isacharparenleft}{\isasymPhi}\ dE{\isacharparenright}\ of\isanewline
\ \ \ \ PropForm\ {\isasymphi}\ {\isasymRightarrow}\ Term\ {\isacharparenleft}THE\ x{\isachardot}\ case\ {\isacharparenleft}{\isasymPhi}\ x{\isacharparenright}\ of\ PropForm\ {\isasympsi}\ {\isasymRightarrow}\ {\isasympsi}\ cw{\isacharparenright}\isanewline
\ \ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ Error\ dE{\isachardoublequoteclose}%
\begin{isamarkuptext}%
that operator; that \isa{{\isacharparenleft}{\isasymlambda}x{\isachardot}{\isasymphi}{\isacharparenright}} returns Term \isa{{\isacharparenleft}THE\ x{\isachardot}\ {\isasymphi}\ x\ cw{\isacharparenright}}, that is the inbuilt THE 
operator is used and evaluation is wrt to the current world cw; moreover, application of that 
is allowed if \isa{{\isacharparenleft}{\isasymPhi}\ sRE{\isacharparenright}} is a PropForm, otherwise Error is passed on for some someRawEntity%
\end{isamarkuptext}%
\isamarkuptrue%
%
\isamarkupsection{Further logical connectives%
}
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ z{\isacharunderscore}and{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}io\ opt{\isasymRightarrow}io\ opt{\isasymRightarrow}io\ opt{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymand}\isactrlsup z\ {\isasympsi}\ {\isasymequiv}\ {\isasymnot}\isactrlsup z{\isacharparenleft}{\isasymphi}\ {\isasymrightarrow}\isactrlsup z\ {\isasymnot}\isactrlsup z{\isasympsi}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ z{\isacharunderscore}or{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}io\ opt{\isasymRightarrow}io\ opt{\isasymRightarrow}io\ opt{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymor}\isactrlsup z\ {\isasympsi}\ {\isasymequiv}\ {\isacharparenleft}{\isasymnot}\isactrlsup z{\isasymphi}\ {\isasymrightarrow}\isactrlsup z\ {\isasympsi}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ z{\isacharunderscore}equiv{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}io\ opt{\isasymRightarrow}io\ opt{\isasymRightarrow}io\ opt{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymphi}\ {\isasymequiv}\isactrlsup z\ {\isasympsi}\ {\isasymequiv}\ {\isacharparenleft}{\isasymphi}\ {\isasymrightarrow}\isactrlsup z\ {\isasympsi}{\isacharparenright}\ {\isasymand}\isactrlsup z\ {\isacharparenleft}{\isasympsi}\ {\isasymrightarrow}\isactrlsup z\ {\isasymphi}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ z{\isacharunderscore}exists{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}{\isacharprime}a{\isasymRightarrow}io\ opt{\isacharparenright}{\isasymRightarrow}io\ opt{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymexists}\ {\isasymPhi}\ {\isasymequiv}\ case\ {\isacharparenleft}{\isasymPhi}\ dA{\isacharparenright}\ of\isanewline
\ \ \ \ PropForm\ {\isasymphi}\ {\isasymRightarrow}\ PropForm\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymexists}x{\isachardot}\ case\ {\isacharparenleft}{\isasymPhi}\ x{\isacharparenright}\ of\ PropForm\ {\isasympsi}\ {\isasymRightarrow}\ {\isasympsi}\ w{\isacharparenright}\isanewline
\ \ {\isacharbar}\ Form\ {\isasymphi}\ {\isasymRightarrow}\ Form\ {\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymexists}x{\isachardot}\ case\ {\isacharparenleft}{\isasymPhi}\ x{\isacharparenright}\ of\ Form\ {\isasympsi}\ {\isasymRightarrow}\ {\isasympsi}\ w{\isacharparenright}\isanewline
\ \ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ Error\ dIO{\isachardoublequoteclose}\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ z{\isacharunderscore}dia{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}io\ opt{\isasymRightarrow}io\ opt{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymdiamond}\ {\isasymphi}\ {\isasymequiv}\ {\isasymnot}\isactrlsup z\ {\isacharparenleft}{\isasymbox}\ {\isacharparenleft}{\isasymnot}\isactrlsup z\ {\isasymphi}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%
\isamarkupsection{Some shortcuts for the constructors%
}
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ mkPropForm\ {\isacharcolon}{\isacharcolon}\ \ {\isachardoublequoteopen}io{\isasymRightarrow}io\ opt{\isachardoublequoteclose}\ \ \isakeyword{where}\ {\isachardoublequoteopen}{\isacharcomma}p{\isacharcomma}\ {\isasymequiv}\ PropForm\ p{\isachardoublequoteclose}\ \isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ mkForm\ {\isacharcolon}{\isacharcolon}\ \ {\isachardoublequoteopen}io{\isasymRightarrow}io\ opt{\isachardoublequoteclose}\ \ \isakeyword{where}\ {\isachardoublequoteopen}{\isacharsemicolon}p{\isacharsemicolon}\ {\isasymequiv}\ Form\ p{\isachardoublequoteclose}\ \isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ mkTerm\ {\isacharcolon}{\isacharcolon}\ \ {\isachardoublequoteopen}{\isacharprime}a{\isasymRightarrow}{\isacharprime}a\ opt{\isachardoublequoteclose}\ \ \isakeyword{where}\ {\isachardoublequoteopen}{\isachardot}t{\isachardot}\ {\isasymequiv}\ Term\ t{\isachardoublequoteclose}%
\isamarkupsection{Some basic tests%
}
\isamarkuptrue%
%
\begin{isamarkuptext}%
Verifying Modal Logic Principles%
\end{isamarkuptext}%
\isamarkuptrue%
%
\begin{isamarkuptext}%
Necessitation holds%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ necessitation{\isacharunderscore}PropForm{\isacharcolon}\ {\isachardoublequoteopen}{\isasymforall}{\isasymphi}{\isachardot}\ {\isacharbrackleft}{\isacharcomma}{\isasymphi}{\isacharcomma}{\isacharbrackright}\ {\isasymlongrightarrow}\ {\isacharbrackleft}{\isasymbox}\ {\isacharcomma}{\isasymphi}{\isacharcomma}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ necessitation{\isacharunderscore}Form{\isacharcolon}\ \ \ \ \ {\isachardoublequoteopen}{\isasymforall}{\isasymphi}{\isachardot}\ {\isacharbrackleft}{\isacharsemicolon}{\isasymphi}{\isacharsemicolon}{\isacharbrackright}\ {\isasymlongrightarrow}\ {\isacharbrackleft}{\isasymbox}\ {\isacharsemicolon}{\isasymphi}{\isacharsemicolon}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
Modal Collapse does not hold%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ modalCollapse{\isacharunderscore}PropForm{\isacharcolon}\ {\isachardoublequoteopen}{\isasymforall}{\isasymphi}{\isachardot}\ {\isacharbrackleft}{\isacharcomma}{\isasymphi}{\isacharcomma}\ {\isasymrightarrow}\isactrlsup z\ {\isasymbox}\ {\isacharcomma}{\isasymphi}{\isacharcomma}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\ \isacommand{nitpick}\isamarkupfalse%
%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{oops}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ modalCollapse{\isacharunderscore}Form{\isacharcolon}\ \ \ \ \ {\isachardoublequoteopen}{\isasymforall}{\isasymphi}{\isachardot}\ {\isacharbrackleft}{\isacharsemicolon}{\isasymphi}{\isacharsemicolon}\ {\isasymrightarrow}\isactrlsup z\ {\isasymbox}\ {\isacharsemicolon}{\isasymphi}{\isacharsemicolon}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\ \isacommand{nitpick}\isamarkupfalse%
%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{oops}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
Verifying S5 Principles%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ axiom{\isacharunderscore}M{\isacharunderscore}PropForm{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}{\isacharparenleft}{\isasymlambda}{\isasymphi}{\isachardot}\ {\isacharparenleft}{\isasymbox}\ {\isacharcomma}{\isasymphi}{\isacharcomma}{\isacharparenright}\ {\isasymrightarrow}\isactrlsup z\ {\isacharcomma}{\isasymphi}{\isacharcomma}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ axiom{\isacharunderscore}M{\isacharunderscore}Form{\isacharcolon}\ \ \ \ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}{\isacharparenleft}{\isasymlambda}{\isasymphi}{\isachardot}\ {\isacharparenleft}{\isasymbox}\ {\isacharsemicolon}{\isasymphi}{\isacharsemicolon}{\isacharparenright}\ {\isasymrightarrow}\isactrlsup z\ {\isacharsemicolon}{\isasymphi}{\isacharsemicolon}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ axiom{\isacharunderscore}B{\isacharunderscore}PropForm{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}{\isacharparenleft}{\isasymlambda}{\isasymphi}{\isachardot}\ {\isacharcomma}{\isasymphi}{\isacharcomma}\ {\isasymrightarrow}\isactrlsup z\ {\isacharparenleft}{\isasymbox}\ {\isacharparenleft}{\isasymdiamond}\ {\isacharcomma}{\isasymphi}{\isacharcomma}{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ axiom{\isacharunderscore}B{\isacharunderscore}Form{\isacharcolon}\ \ \ \ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}{\isacharparenleft}{\isasymlambda}{\isasymphi}{\isachardot}\ {\isacharsemicolon}{\isasymphi}{\isacharsemicolon}\ {\isasymrightarrow}\isactrlsup z\ {\isacharparenleft}{\isasymbox}\ {\isacharparenleft}{\isasymdiamond}\ {\isacharsemicolon}{\isasymphi}{\isacharsemicolon}{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ axiom{\isacharunderscore}D{\isacharunderscore}PropForm{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}{\isacharparenleft}{\isasymlambda}{\isasymphi}{\isachardot}\ {\isacharparenleft}{\isasymbox}\ {\isacharcomma}{\isasymphi}{\isacharcomma}{\isacharparenright}\ {\isasymrightarrow}\isactrlsup z\ {\isacharparenleft}{\isasymbox}\ {\isacharparenleft}{\isasymbox}\ {\isacharcomma}{\isasymphi}{\isacharcomma}{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ axiom{\isacharunderscore}D{\isacharunderscore}Form{\isacharcolon}\ \ \ \ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}{\isacharparenleft}{\isasymlambda}{\isasymphi}{\isachardot}\ {\isacharparenleft}{\isasymbox}\ {\isacharsemicolon}{\isasymphi}{\isacharsemicolon}{\isacharparenright}\ {\isasymrightarrow}\isactrlsup z\ {\isacharparenleft}{\isasymbox}\ {\isacharparenleft}{\isasymbox}\ {\isacharsemicolon}{\isasymphi}{\isacharsemicolon}{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ axiom{\isacharunderscore}{\isadigit{4}}{\isacharunderscore}PropForm{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}{\isacharparenleft}{\isasymlambda}{\isasymphi}{\isachardot}\ {\isacharparenleft}{\isasymbox}\ {\isacharcomma}{\isasymphi}{\isacharcomma}{\isacharparenright}\ {\isasymrightarrow}\isactrlsup z\ {\isacharparenleft}{\isasymdiamond}\ {\isacharcomma}{\isasymphi}{\isacharcomma}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ axiom{\isacharunderscore}{\isadigit{4}}{\isacharunderscore}Form{\isacharcolon}\ \ \ \ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}{\isacharparenleft}{\isasymlambda}{\isasymphi}{\isachardot}\ {\isacharparenleft}{\isasymbox}\ {\isacharsemicolon}{\isasymphi}{\isacharsemicolon}{\isacharparenright}\ {\isasymrightarrow}\isactrlsup z\ {\isacharparenleft}{\isasymdiamond}\ {\isacharsemicolon}{\isasymphi}{\isacharsemicolon}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ axiom{\isacharunderscore}{\isadigit{5}}{\isacharunderscore}PropForm{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}{\isacharparenleft}{\isasymlambda}{\isasymphi}{\isachardot}\ {\isacharparenleft}{\isasymdiamond}\ {\isacharcomma}{\isasymphi}{\isacharcomma}{\isacharparenright}\ {\isasymrightarrow}\isactrlsup z\ {\isacharparenleft}{\isasymbox}\ {\isacharparenleft}{\isasymdiamond}\ {\isacharcomma}{\isasymphi}{\isacharcomma}{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ axiom{\isacharunderscore}{\isadigit{5}}{\isacharunderscore}Form{\isacharcolon}\ \ \ \ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}{\isacharparenleft}{\isasymlambda}{\isasymphi}{\isachardot}\ {\isacharparenleft}{\isasymdiamond}\ {\isacharsemicolon}{\isasymphi}{\isacharsemicolon}{\isacharparenright}\ {\isasymrightarrow}\isactrlsup z\ {\isacharparenleft}{\isasymbox}\ {\isacharparenleft}{\isasymdiamond}\ {\isacharsemicolon}{\isasymphi}{\isacharsemicolon}{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ test{\isacharunderscore}A{\isacharunderscore}PropForm{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}{\isacharparenleft}{\isasymlambda}{\isasymphi}{\isachardot}\ {\isacharparenleft}{\isasymbox}\ {\isacharparenleft}{\isasymdiamond}\ {\isacharcomma}{\isasymphi}{\isacharcomma}{\isacharparenright}{\isacharparenright}\ {\isasymrightarrow}\isactrlsup z\ {\isacharparenleft}{\isasymdiamond}\ {\isacharcomma}{\isasymphi}{\isacharcomma}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ test{\isacharunderscore}A{\isacharunderscore}Form{\isacharcolon}\ \ \ \ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}{\isacharparenleft}{\isasymlambda}{\isasymphi}{\isachardot}\ {\isacharparenleft}{\isasymbox}\ {\isacharparenleft}{\isasymdiamond}\ {\isacharsemicolon}{\isasymphi}{\isacharsemicolon}{\isacharparenright}{\isacharparenright}\ {\isasymrightarrow}\isactrlsup z\ {\isacharparenleft}{\isasymdiamond}\ {\isacharsemicolon}{\isasymphi}{\isacharsemicolon}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ test{\isacharunderscore}B{\isacharunderscore}PropForm{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}{\isacharparenleft}{\isasymlambda}{\isasymphi}{\isachardot}\ {\isacharparenleft}{\isasymdiamond}\ {\isacharparenleft}{\isasymbox}\ {\isacharcomma}{\isasymphi}{\isacharcomma}{\isacharparenright}{\isacharparenright}\ {\isasymrightarrow}\isactrlsup z\ {\isacharparenleft}{\isasymdiamond}\ {\isacharcomma}{\isasymphi}{\isacharcomma}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ test{\isacharunderscore}B{\isacharunderscore}Form{\isacharcolon}\ \ \ \ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}{\isacharparenleft}{\isasymlambda}{\isasymphi}{\isachardot}\ {\isacharparenleft}{\isasymdiamond}\ {\isacharparenleft}{\isasymbox}\ {\isacharsemicolon}{\isasymphi}{\isacharsemicolon}{\isacharparenright}{\isacharparenright}\ {\isasymrightarrow}\isactrlsup z\ {\isacharparenleft}{\isasymdiamond}\ {\isacharsemicolon}{\isasymphi}{\isacharsemicolon}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ test{\isacharunderscore}C{\isacharunderscore}PropForm{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}{\isacharparenleft}{\isasymlambda}{\isasymphi}{\isachardot}\ {\isacharparenleft}{\isasymdiamond}\ {\isacharparenleft}{\isasymbox}\ {\isacharcomma}{\isasymphi}{\isacharcomma}{\isacharparenright}{\isacharparenright}\ {\isasymrightarrow}\isactrlsup z\ {\isacharparenleft}{\isasymdiamond}\ {\isacharcomma}{\isasymphi}{\isacharcomma}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{by}\isamarkupfalse%
\ metis%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ test{\isacharunderscore}C{\isacharunderscore}Form{\isacharcolon}\ \ \ \ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}{\isacharparenleft}{\isasymlambda}{\isasymphi}{\isachardot}\ {\isacharparenleft}{\isasymdiamond}\ {\isacharparenleft}{\isasymbox}\ {\isacharsemicolon}{\isasymphi}{\isacharsemicolon}{\isacharparenright}{\isacharparenright}\ {\isasymrightarrow}\isactrlsup z\ {\isacharparenleft}{\isasymdiamond}\ {\isacharsemicolon}{\isasymphi}{\isacharsemicolon}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{by}\isamarkupfalse%
\ metis%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ test{\isacharunderscore}D{\isacharunderscore}PropForm{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}{\isacharparenleft}{\isasymlambda}{\isasymphi}{\isachardot}\ {\isacharparenleft}{\isasymdiamond}\ {\isacharparenleft}{\isasymbox}\ {\isacharcomma}{\isasymphi}{\isacharcomma}{\isacharparenright}{\isacharparenright}\ {\isasymrightarrow}\isactrlsup z\ {\isacharparenleft}{\isasymbox}\ {\isacharcomma}{\isasymphi}{\isacharcomma}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ test{\isacharunderscore}D{\isacharunderscore}Form{\isacharcolon}\ \ \ \ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}{\isacharparenleft}{\isasymlambda}{\isasymphi}{\isachardot}\ {\isacharparenleft}{\isasymdiamond}\ {\isacharparenleft}{\isasymbox}\ {\isacharsemicolon}{\isasymphi}{\isacharsemicolon}{\isacharparenright}{\isacharparenright}\ {\isasymrightarrow}\isactrlsup z\ {\isacharparenleft}{\isasymbox}\ {\isacharsemicolon}{\isasymphi}{\isacharsemicolon}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
Example signature; entities and relations%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{consts}\isamarkupfalse%
\ a{\isacharunderscore}{\isadigit{0}}\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}e{\isachardoublequoteclose}\ \isacommand{abbreviation}\isamarkupfalse%
\ a\ \ \isakeyword{where}\ {\isachardoublequoteopen}a\ {\isasymequiv}\ {\isachardot}a{\isacharunderscore}{\isadigit{0}}{\isachardot}{\isachardoublequoteclose}\isanewline
\isacommand{consts}\isamarkupfalse%
\ b{\isacharunderscore}{\isadigit{0}}\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}e{\isachardoublequoteclose}\ \isacommand{abbreviation}\isamarkupfalse%
\ b\ \ \isakeyword{where}\ {\isachardoublequoteopen}b\ {\isasymequiv}\ {\isachardot}b{\isacharunderscore}{\isadigit{0}}{\isachardot}{\isachardoublequoteclose}\isanewline
\isacommand{consts}\isamarkupfalse%
\ c{\isacharunderscore}{\isadigit{0}}\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}e{\isachardoublequoteclose}\ \isacommand{abbreviation}\isamarkupfalse%
\ c\ \ \isakeyword{where}\ {\isachardoublequoteopen}c\ {\isasymequiv}\ {\isachardot}c{\isacharunderscore}{\isadigit{0}}{\isachardot}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{consts}\isamarkupfalse%
\ R{\isacharunderscore}{\isadigit{0}}\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}io{\isachardoublequoteclose}\ \ \isacommand{abbreviation}\isamarkupfalse%
\ R{\isadigit{0}}\ \ \isakeyword{where}\ {\isachardoublequoteopen}R{\isadigit{0}}\ {\isasymequiv}\ {\isachardot}R{\isacharunderscore}{\isadigit{0}}{\isachardot}{\isachardoublequoteclose}\isanewline
\isacommand{consts}\isamarkupfalse%
\ R{\isacharunderscore}{\isadigit{1}}\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}e{\isasymRightarrow}io{\isachardoublequoteclose}\ \isacommand{abbreviation}\isamarkupfalse%
\ R{\isadigit{1}}\ \ \isakeyword{where}\ {\isachardoublequoteopen}R{\isadigit{1}}\ {\isasymequiv}\ {\isachardot}R{\isacharunderscore}{\isadigit{1}}{\isachardot}{\isachardoublequoteclose}\isanewline
\isacommand{consts}\isamarkupfalse%
\ R{\isacharunderscore}{\isadigit{2}}\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}e{\isasymRightarrow}e{\isasymRightarrow}io{\isachardoublequoteclose}\ \isacommand{abbreviation}\isamarkupfalse%
\ R{\isadigit{2}}\ \ \isakeyword{where}\ {\isachardoublequoteopen}R{\isadigit{2}}\ {\isasymequiv}\ {\isachardot}R{\isacharunderscore}{\isadigit{2}}{\isachardot}{\isachardoublequoteclose}\isanewline
\isacommand{consts}\isamarkupfalse%
\ R{\isacharunderscore}{\isadigit{3}}\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}e{\isasymRightarrow}e{\isasymRightarrow}e{\isasymRightarrow}io{\isachardoublequoteclose}\ \ \isacommand{abbreviation}\isamarkupfalse%
\ R{\isadigit{3}}\ \ \isakeyword{where}\ {\isachardoublequoteopen}R{\isadigit{3}}\ {\isasymequiv}\ {\isachardot}R{\isacharunderscore}{\isadigit{3}}{\isachardot}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
Testing term and formula constructions%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharless}R{\isadigit{1}}{\isasymbullet}a{\isachargreater}{\isacharbrackright}{\isachardoublequoteclose}\ \isacommand{nitpick}\isamarkupfalse%
%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{oops}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}isPropForm\ {\isacharless}R{\isadigit{1}}{\isasymbullet}a{\isachargreater}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}R{\isadigit{1}}{\isasymbullet}a{\isachargreater}\ {\isacharequal}\ X{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{oops}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharless}a{\isasymcirc}R{\isadigit{1}}{\isachargreater}{\isacharbrackright}{\isachardoublequoteclose}\ \isacommand{nitpick}\isamarkupfalse%
%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{oops}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}isPropForm\ {\isacharless}a{\isasymcirc}R{\isadigit{1}}{\isachargreater}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{oops}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}isForm\ {\isacharless}a{\isasymcirc}R{\isadigit{1}}{\isachargreater}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}a{\isasymcirc}R{\isadigit{1}}{\isachargreater}\ {\isacharequal}\ X{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{oops}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharless}{\isasymlambda}\isactrlsup {\isadigit{1}}{\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isacharless}R{\isadigit{1}}{\isasymbullet}{\isachardot}x{\isachardot}{\isachargreater}\ {\isasymrightarrow}\isactrlsup z\ {\isacharless}R{\isadigit{1}}{\isasymbullet}{\isachardot}x{\isachardot}{\isachargreater}{\isacharparenright}{\isasymbullet}a{\isachargreater}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\isasymlambda}\isactrlsup {\isadigit{1}}{\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isacharless}R{\isadigit{1}}{\isasymbullet}{\isachardot}x{\isachardot}{\isachargreater}\ {\isasymrightarrow}\isactrlsup z\ {\isacharless}R{\isadigit{1}}{\isasymbullet}{\isachardot}x{\isachardot}{\isachargreater}{\isacharparenright}{\isasymbullet}a{\isachargreater}\ {\isacharequal}\ X{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{oops}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymnot}\ isWff\ {\isacharparenleft}{\isacharless}R{\isadigit{1}}{\isasymbullet}{\isachardot}x{\isachardot}{\isachargreater}\ {\isasymrightarrow}\isactrlsup z\ {\isacharless}{\isachardot}x{\isachardot}{\isasymcirc}R{\isadigit{1}}{\isachargreater}{\isacharparenright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlambda}\isactrlsup {\isadigit{1}}{\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isacharless}R{\isadigit{1}}{\isasymbullet}{\isachardot}x{\isachardot}{\isachargreater}\ {\isasymrightarrow}\isactrlsup z\ {\isacharless}{\isachardot}x{\isachardot}{\isasymcirc}R{\isadigit{1}}{\isachargreater}{\isacharparenright}\ {\isacharequal}\ X{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{oops}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharless}{\isasymlambda}\isactrlsup {\isadigit{1}}{\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isacharless}R{\isadigit{1}}{\isasymbullet}{\isachardot}x{\isachardot}{\isachargreater}\ {\isasymrightarrow}\isactrlsup z\ {\isacharless}{\isachardot}x{\isachardot}{\isasymcirc}R{\isadigit{1}}{\isachargreater}{\isacharparenright}{\isasymbullet}a{\isachargreater}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{oops}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharless}{\isasymlambda}\isactrlsup {\isadigit{1}}{\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isacharless}R{\isadigit{1}}{\isasymbullet}{\isachardot}x{\isachardot}{\isachargreater}\ {\isasymrightarrow}\isactrlsup z\ {\isacharless}{\isachardot}x{\isachardot}{\isasymcirc}R{\isadigit{1}}{\isachargreater}{\isacharparenright}{\isasymbullet}a{\isachargreater}\ {\isacharequal}\ X{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{oops}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}{\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isacharless}R{\isadigit{1}}{\isasymbullet}{\isachardot}x{\isachardot}{\isachargreater}\ {\isasymrightarrow}\isactrlsup z\ {\isacharless}R{\isadigit{1}}{\isasymbullet}{\isachardot}x{\isachardot}{\isachargreater}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}{\isacharparenleft}{\isasymlambda}R{\isachardot}\ {\isasymforall}{\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isacharless}{\isachardot}R{\isachardot}{\isasymbullet}{\isachardot}x{\isachardot}{\isachargreater}\ {\isasymrightarrow}\isactrlsup z\ {\isacharless}{\isachardot}R{\isachardot}{\isasymbullet}{\isachardot}x{\isachardot}{\isachargreater}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}{\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isacharless}R{\isadigit{1}}{\isasymbullet}{\isachardot}x{\isachardot}{\isachargreater}\ {\isasymrightarrow}\isactrlsup z\ {\isacharless}R{\isadigit{1}}{\isasymbullet}{\isachardot}x{\isachardot}{\isachargreater}{\isacharparenright}\ {\isacharequal}\ X{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{oops}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}{\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isacharless}{\isachardot}x{\isachardot}{\isasymcirc}R{\isadigit{1}}{\isachargreater}\ {\isasymrightarrow}\isactrlsup z\ {\isacharless}{\isachardot}x{\isachardot}{\isasymcirc}R{\isadigit{1}}{\isachargreater}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}{\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isacharless}{\isachardot}x{\isachardot}{\isasymcirc}R{\isadigit{1}}{\isachargreater}\ {\isasymrightarrow}\isactrlsup z\ {\isacharless}{\isachardot}x{\isachardot}{\isasymcirc}R{\isadigit{1}}{\isachargreater}{\isacharparenright}\ {\isacharequal}\ X{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{oops}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}{\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isacharless}R{\isadigit{1}}{\isasymbullet}{\isachardot}x{\isachardot}{\isachargreater}\ {\isasymrightarrow}\isactrlsup z\ {\isacharless}{\isachardot}x{\isachardot}{\isasymcirc}R{\isadigit{1}}{\isachargreater}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{oops}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}{\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isacharless}R{\isadigit{1}}{\isasymbullet}{\isachardot}x{\isachardot}{\isachargreater}\ {\isasymrightarrow}\isactrlsup z\ {\isacharless}{\isachardot}x{\isachardot}{\isasymcirc}R{\isadigit{1}}{\isachargreater}{\isacharparenright}\ {\isacharequal}\ X{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{oops}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}{\isacharparenleft}{\isasymlambda}R{\isachardot}\ {\isacharless}{\isachardot}R{\isachardot}{\isasymbullet}{\isachardot}x{\isachardot}{\isachargreater}\ {\isasymrightarrow}\isactrlsup z\ {\isacharless}{\isachardot}x{\isachardot}{\isasymcirc}{\isachardot}R{\isachardot}{\isachargreater}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{oops}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymforall}{\isacharparenleft}{\isasymlambda}R{\isachardot}\ {\isacharless}{\isachardot}R{\isachardot}{\isasymbullet}{\isachardot}x{\isachardot}{\isachargreater}\ {\isasymrightarrow}\isactrlsup z\ {\isacharless}{\isachardot}x{\isachardot}{\isasymcirc}{\isachardot}R{\isachardot}{\isachargreater}{\isacharparenright}\ {\isacharequal}\ X{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{oops}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\isamarkupsection{Are the priorities set correctly?%
}
\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharcomma}{\isasymphi}{\isacharcomma}\ {\isasymand}\isactrlsup z\ {\isacharcomma}{\isasympsi}{\isacharcomma}\ {\isasymrightarrow}\isactrlsup z\ {\isacharcomma}{\isasymchi}{\isacharcomma}\ {\isasymequiv}\ {\isacharparenleft}{\isacharcomma}{\isasymphi}{\isacharcomma}\ {\isasymand}\isactrlsup z\ {\isacharcomma}{\isasympsi}{\isacharcomma}{\isacharparenright}\ {\isasymrightarrow}\isactrlsup z\ {\isacharcomma}{\isasymchi}{\isacharcomma}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharcomma}{\isasymphi}{\isacharcomma}\ {\isasymand}\isactrlsup z\ {\isacharcomma}{\isasympsi}{\isacharcomma}\ {\isasymrightarrow}\isactrlsup z\ {\isacharcomma}{\isasymchi}{\isacharcomma}\ {\isasymequiv}\ {\isacharcomma}{\isasymphi}{\isacharcomma}\ {\isasymand}\isactrlsup z\ {\isacharparenleft}{\isacharcomma}{\isasympsi}{\isacharcomma}\ {\isasymrightarrow}\isactrlsup z\ {\isacharcomma}{\isasymchi}{\isacharcomma}{\isacharparenright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\ \isacommand{nitpick}\isamarkupfalse%
%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{oops}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\isacharcomma}{\isasymphi}{\isacharcomma}\ {\isasymand}\isactrlsup z\ {\isacharcomma}{\isasympsi}{\isacharcomma}\ {\isasymequiv}\isactrlsup z\ {\isacharcomma}{\isasymphi}{\isacharcomma}\ {\isasymand}\isactrlsup z\ {\isacharcomma}{\isasympsi}{\isacharcomma}{\isacharparenright}\ {\isasymequiv}\ {\isacharparenleft}{\isacharparenleft}{\isacharcomma}{\isasymphi}{\isacharcomma}\ {\isasymand}\isactrlsup z\ {\isacharcomma}{\isasympsi}{\isacharcomma}{\isacharparenright}\ {\isasymequiv}\isactrlsup z\ {\isacharparenleft}{\isacharcomma}{\isasymphi}{\isacharcomma}\ {\isasymand}\isactrlsup z\ {\isacharcomma}{\isasympsi}{\isacharcomma}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\isacharcomma}{\isasymphi}{\isacharcomma}\ {\isasymand}\isactrlsup z\ {\isacharcomma}{\isasympsi}{\isacharcomma}\ {\isasymequiv}\isactrlsup z\ {\isacharcomma}{\isasymphi}{\isacharcomma}\ {\isasymand}\isactrlsup z\ {\isacharcomma}{\isasympsi}{\isacharcomma}{\isacharparenright}\ {\isasymequiv}\ {\isacharparenleft}{\isacharcomma}{\isasymphi}{\isacharcomma}\ {\isasymand}\isactrlsup z\ {\isacharparenleft}{\isacharcomma}{\isasympsi}{\isacharcomma}\ {\isasymequiv}\isactrlsup z\ {\isacharcomma}{\isasymphi}{\isacharcomma}{\isacharparenright}\ {\isasymand}\isactrlsup z\ {\isacharcomma}{\isasympsi}{\isacharcomma}{\isacharparenright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\ \isacommand{nitpick}\isamarkupfalse%
%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{oops}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\isamarkupsection{E!, O!, A! and =E%
}
\isamarkuptrue%
\isacommand{consts}\isamarkupfalse%
\ E{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}e{\isasymRightarrow}io{\isacharparenright}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
Distinguished 1-place relation constant: E! (read: ‘being concrete’ or ‘concreteness’)%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ z{\isacharunderscore}ordinary{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}e{\isasymRightarrow}io{\isacharparenright}\ opt{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}O\isactrlsup {\isacharbang}\ {\isasymequiv}\ {\isasymlambda}\isactrlsup {\isadigit{1}}{\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isasymdiamond}\ {\isacharless}{\isachardot}E{\isachardot}{\isasymbullet}{\isachardot}x{\isachardot}{\isachargreater}{\isacharparenright}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
Being ordinary is being possibly concrete.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ z{\isacharunderscore}abstract{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}e{\isasymRightarrow}io{\isacharparenright}\ opt{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}A\isactrlsup {\isacharbang}\ {\isasymequiv}\ {\isasymlambda}\isactrlsup {\isadigit{1}}{\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isasymnot}\isactrlsup z\ {\isacharparenleft}{\isasymdiamond}\ {\isacharless}{\isachardot}E{\isachardot}{\isasymbullet}{\isachardot}x{\isachardot}{\isachargreater}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
Being abstract is not possibly being concrete.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ z{\isacharunderscore}identity{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}e{\isasymRightarrow}e{\isasymRightarrow}io{\isacharparenright}\ opt{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isacharequal}\isactrlsub e\isactrlsup z\ {\isasymequiv}\ \isanewline
\ \ {\isasymlambda}\isactrlsup {\isadigit{2}}{\isacharparenleft}{\isasymlambda}x\ y{\isachardot}\ {\isacharparenleft}{\isacharparenleft}{\isacharless}O\isactrlsup {\isacharbang}{\isasymbullet}{\isachardot}x{\isachardot}{\isachargreater}\ {\isasymand}\isactrlsup z\ {\isacharless}O\isactrlsup {\isacharbang}{\isasymbullet}{\isachardot}y{\isachardot}{\isachargreater}{\isacharparenright}\ {\isasymand}\isactrlsup z\ {\isasymbox}\ {\isacharparenleft}{\isasymforall}{\isacharparenleft}{\isasymlambda}F{\isachardot}\ {\isacharless}{\isachardot}F{\isachardot}{\isasymbullet}{\isachardot}x{\isachardot}{\isachargreater}\ {\isasymequiv}\isactrlsup z\ {\isacharless}{\isachardot}F{\isachardot}{\isasymbullet}{\isachardot}y{\isachardot}{\isachargreater}{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ z{\isacharunderscore}identityE{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}e\ opt{\isasymRightarrow}e\ opt{\isasymRightarrow}io\ opt{\isacharparenright}{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}x\ {\isacharequal}\isactrlsub E\ y\ {\isasymequiv}\ {\isacharparenleft}Exe{\isadigit{2}}\ {\isacharequal}\isactrlsub e\isactrlsup z\ x\ y{\isacharparenright}{\isachardoublequoteclose}%
\isamarkupsection{Further test examples%
}
\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}{\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isasymexists}{\isacharparenleft}{\isasymlambda}R{\isachardot}\ {\isacharparenleft}{\isacharless}{\isachardot}x{\isachardot}{\isasymcirc}{\isachardot}R{\isachardot}{\isachargreater}\ {\isasymrightarrow}\isactrlsup z\ {\isacharless}{\isachardot}x{\isachardot}{\isasymcirc}R{\isadigit{1}}{\isachargreater}{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}{\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isasymforall}{\isacharparenleft}{\isasymlambda}R{\isachardot}\ {\isacharparenleft}{\isacharless}{\isachardot}x{\isachardot}{\isasymcirc}{\isachardot}R{\isachardot}{\isachargreater}\ {\isasymrightarrow}\isactrlsup z\ {\isacharless}{\isachardot}x{\isachardot}{\isasymcirc}R{\isadigit{1}}{\isachargreater}{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\ \isacommand{nitpick}\isamarkupfalse%
%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{oops}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}a\ {\isacharequal}\isactrlsub E\ a{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\ \isacommand{nitpick}\isamarkupfalse%
%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{oops}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharless}O\isactrlsup {\isacharbang}{\isasymbullet}a{\isachargreater}\ {\isasymrightarrow}\isactrlsup z\ a\ {\isacharequal}\isactrlsub E\ a{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharparenleft}{\isasymforall}{\isacharparenleft}{\isasymlambda}F{\isachardot}\ {\isacharless}{\isachardot}F{\isachardot}{\isasymbullet}{\isachardot}x{\isachardot}{\isachargreater}\ {\isasymequiv}\isactrlsup z\ {\isacharless}{\isachardot}F{\isachardot}{\isasymbullet}{\isachardot}x{\isachardot}{\isachargreater}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharless}O\isactrlsup {\isacharbang}{\isasymbullet}a{\isachargreater}\ {\isasymrightarrow}\isactrlsup z\ {\isacharless}{\isasymlambda}\isactrlsup {\isadigit{1}}{\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isachardot}x{\isachardot}\ {\isacharequal}\isactrlsub E\ a{\isacharparenright}{\isasymbullet}a{\isachargreater}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharparenleft}{\isasymexists}{\isacharparenleft}{\isasymlambda}F{\isachardot}\ {\isacharless}a{\isasymcirc}{\isachardot}F{\isachardot}{\isachargreater}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}isWff\ {\isacharcomma}{\isacharparenleft}{\isasymlambda}w{\isachardot}\ True{\isacharparenright}{\isacharcomma}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharparenleft}{\isasymexists}{\isacharparenleft}{\isasymlambda}F{\isachardot}\ {\isacharcomma}F{\isacharcomma}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharparenleft}{\isasymexists}{\isacharparenleft}{\isasymlambda}F{\isachardot}\ {\isacharsemicolon}F{\isacharsemicolon}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ {\isacharparenleft}simp{\isacharparenright}\ \isacommand{by}\isamarkupfalse%
\ auto\isanewline
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
\end{isabellebody}%
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "root"
%%% End: