### 2015-12-10 PrincipiaMetaphysicaAutomationPart1/doc/document/TheoryOfAbstractObjectsV1.tex
latest syntax improvements: binder notation and various improvements

%
\begin{isabellebody}%
\def\isabellecontext{TheoryOfAbstractObjectsV{\isadigit{1}}}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isamarkupsection{Introduction%
}
\isamarkuptrue%
%
\begin{isamarkuptext}%
We present a formalisation and partial automation of an initial part of the (third authors)
 Principia Metaphysica \cite{zalta:_princ_metap} in Isabelle/HOL \cite{Isabelle}.

 The Principia Metaphysica, which is based on and extends the Theory of Abstract 
 Objects \cite{zalta83:_abstr_objec}, 
 employs a modal relational type theory as logical foundation. Arguments
 defending this choice against a modal functional type theory 
 have been presented before \cite{zalta11:_relat_versus_funct_found_logic}.
 In a nutshell, the situation is this: functional type theory comes with strong 
 comprehension principles, which, in the context of the Theory of Abstract Objects, 
 have paradoxical implications. When starting off with a relational foundation, however, 
 weaker comprehension principles are provided, and these obstacles can be avoided.

 Isabelle/HOL is a proof assistant based on a functional type theory, 
 more precisely, Church's theory of types \cite{Church40}. Recently, it has been shown 
 that Church's type theory can be elegantly utilized as a meta-logic to encode and 
 automate various quantified non-classical logics, including modal functional type 
 theory \cite{J23,C40}. This work has subsequently been employed in a case study in
 computational metaphysics, in which different variants of Kurt Gödel's ontological 
 argument \cite{ECAI} were verified (respectively, falsified).

 The motivating research questions for the work presented below include:
 \begin{itemize} 
 \item Can functional type theory, despite the problems as pointed by Zalta and Oppenheimer, 
  be utilized to encode the Theory of Abstract Objects when following the embeddings approach?
 \item How elegant and user-friendly is the resulting formalization? In other
  words, to what extend can Isabelle's  user interface be facilitated to hide 
  unpleasant technicalities of the (extended) embedding from the user?
 \item How far can automation be pushed in the approach? How much user interaction can
  be avoided in the formalization of the (first part) of the Principia Metaphysica? 
 \item Can the consistency of the theory be validated with the available automated 
  reasoning tools?
 \item Can the reasoners eventually even contribute some new knowledge? 
 \ldots
 \end{itemize}%
\end{isamarkuptext}%
\isamarkuptrue%
%
\begin{isamarkuptext}%
The encoding of modal functional type theory in functional type theory as explored in 
 previous work \cite{J23,C40} is simple: modal logic formulas are identified with certain functional 
 type theory formulas of predicate type \isa{io}. Possible worlds are explicitly represented by 
 terms of type  \isa{i}. A modal logic \isa{{\isasymphi}} formula holds for a world \isa{w} if and 
 only if the application \isa{{\isasymphi}\ w} evaluates to true. The definition of the propositional modal logic 
 connectives is then straightforward and it simply realizes the standard translation as a set of equations 
 in functional type theory. The approach has been successfully extended for quantifiers. A crucial 
 aspect thereby is that in simple type theory quantifiers can be treated
 as ordinary logical connectives. No extra binding mechanism is needed since the already existing 
 lambda binding mechanism can be elegantly utilized.

 The challenge in this work is to appropriately 'restrict' this embedding for modal relational type theory.
  \begin{figure}[t]
 \includegraphics[height=5.5cm]{ModalRelationalTypeTheory.png}\includegraphics[height=4.5cm]{ModalRelationalTypeTheory2.png}
 \caption{Grammar of Modal Relational Type Theory. \label{mmrt}
 Note that two kinds of (complex) formulas are introduced: ones that may have encoding subformulas and ones that don’t.
 The latter are designated as propositional formulas, the former ones simply as formulas. }
 \end{figure}

 To achieve this  we provide means to explicitly represents and maintain information and constraints on the 
 syntactical structure of modal relational type theory, in particular, we provide means to distinguish 
 between propositional formulas, formulas, terms and erreneous (disallowed) formations. 
 This clearly creates some technical overhead. However, we exploit facilities in Isabelle/HOL's user 
 interface, and other means, to hide most of these technicalities from the user in applications.%
\end{isamarkuptext}%
\isamarkuptrue%
%
\isamarkupsection{Preliminaries%
}
\isamarkuptrue%
%
\begin{isamarkuptext}%
We start out with some type declarations and type abbreviations. 
 Our formalism explicitly encodes Kripke style semantics. Hence, we introduce a 
 distinguished type \isa{i} to represent the set of possible worlds. 
 Consequently, terms of this type denote possible worlds. 
 Moreover, modal logic formulas are associated in our approach with
 predicates (resp. sets) on possible worlds. Hence, modal logic formulas have
 type \isa{{\isacharparenleft}i\ {\isasymRightarrow}\ bool{\isacharparenright}}. To make our representation in the remainder more concise
 we abbreviate this type as \isa{io}.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{typedecl}\isamarkupfalse%
\ i\isanewline
\isacommand{type{\isacharunderscore}synonym}\isamarkupfalse%
\ io\ {\isacharequal}\ {\isachardoublequoteopen}{\isacharparenleft}i\ {\isasymRightarrow}\ bool{\isacharparenright}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
Entities in the abstract theory of types are represented in our formalism by the
 type \isa{e}. We call this the raw type of entities resp. objects. Later 
 on we will introduce means to distinguish between abstract and ordinary entities.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{typedecl}\isamarkupfalse%
\ e%
\begin{isamarkuptext}%
To explicitly model the syntactical restrictions of modal relational type theory we introduce a 
 datatype \isa{{\isacharprime}a\ opt} based on four constructors:
 \isa{ERR\ {\isacharprime}a} (identifies erroneous term constructions), \isa{P\ {\isacharprime}a} (identifies 
 propositional formulas), \isa{F\ {\isacharprime}a} (identifies  formulas), and \isa{T\ {\isacharprime}a} (identifies 
 terms, such as lambda abstractions). The embeddings approach will be suitably adapted below so that 
 for each language expression (in the embedded modal relational type theory) the respective datatype 
 is identified and appropriately propagated. The encapsulated expressions (the polymorphic type \isa{{\isacharprime}a} 
 will be instantiated below) realize the actual modeling of the logic embedding analogous 
 to previous work for modal functional type theory.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{datatype}\isamarkupfalse%
\ {\isacharprime}a\ opt\ {\isacharequal}\ ERR\ {\isacharprime}a\ {\isacharbar}\ P\ {\isacharprime}a\ {\isacharbar}\ F\ {\isacharprime}a\ {\isacharbar}\ T\ {\isacharprime}a%
\begin{isamarkuptext}%
The following operators support a concise and elegant superscript annotation with these
 four syntactical categories for our language constructs.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ mkP{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}io{\isasymRightarrow}io\ opt{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isacharunderscore}\isactrlsup P{\isachardoublequoteclose}\ {\isadigit{1}}{\isadigit{1}}{\isadigit{0}}{\isacharparenright}\ \ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymphi}\isactrlsup P\ {\isasymequiv}\ P\ {\isasymphi}{\isachardoublequoteclose}\ \isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ mkF{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}io{\isasymRightarrow}io\ opt{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isacharunderscore}\isactrlsup F{\isachardoublequoteclose}\ {\isadigit{1}}{\isadigit{1}}{\isadigit{0}}{\isacharparenright}\ \ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymphi}\isactrlsup F\ {\isasymequiv}\ F\ {\isasymphi}{\isachardoublequoteclose}\ \isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ mkT{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharprime}a{\isasymRightarrow}{\isacharprime}a\ opt{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isacharunderscore}\isactrlsup T{\isachardoublequoteclose}\ {\isadigit{1}}{\isadigit{1}}{\isadigit{0}}{\isacharparenright}\ \ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymphi}\isactrlsup T\ {\isasymequiv}\ T\ {\isasymphi}{\isachardoublequoteclose}\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ mkE{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharprime}a{\isasymRightarrow}{\isacharprime}a\ opt{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isacharunderscore}\isactrlsup E{\isachardoublequoteclose}\ {\isadigit{1}}{\isadigit{1}}{\isadigit{0}}{\isacharparenright}\ \ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymphi}\isactrlsup E\ {\isasymequiv}\ ERR\ {\isasymphi}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
Some language constructs in the theory of abstract types, e.g. the actuality operator  
 \isa{{\isasymA}} ("it is actually the case that"), refer to a (fixed) given world. To model such a 
 global world reference we introduce a
 constant symbol (name) \isa{cw} of world type \isa{i}. Moreover, for technical reasons, 
 which will be clarified below, we introduce further (dummy) constant symbols for various domains. Since
 we assume that all domains are non-empty, introducing these constant symbols is obviously not harmful.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{consts}\isamarkupfalse%
\ cw\ {\isacharcolon}{\isacharcolon}\ i\ \isanewline
\isacommand{consts}\isamarkupfalse%
\ de{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}e{\isachardoublequoteclose}\ dio{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}io{\isachardoublequoteclose}\ da{\isacharcolon}{\isacharcolon}{\isacharprime}a%
\isamarkupsection{Embedding of Modal Relational Type Theory%
}
\isamarkuptrue%
%
\begin{isamarkuptext}%
The language constructs of modal relational type theory are introduced step by step.%
\end{isamarkuptext}%
\isamarkuptrue%
%
\begin{isamarkuptext}%
The actuality operator \isa{{\isasymA}} when applied to a formula or propositional formula 
 \isa{{\isasymphi}} evaluates \isa{{\isasymphi}} wrt the fixed given world \isa{cw}. 
 The compound expression \isa{{\isasymA}\ {\isasymphi}} inherits its syntactical category  (\isa{Form} or
 \isa{PropForm}) from \isa{{\isasymphi}}. If the syntactical catagory of  \isa{{\isasymphi}} is 
 \isa{Error} or \isa{Term}, then the syntactical catagory of \isa{{\isasymA}\ {\isasymphi}} 
 is \isa{Error} and a dummy entity of appropriate type is returned. This illustrates the very 
 idea of our explicit structure and constraints and this scheme will repeated below
 for all the other language constructs of modal relational type theory.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ {\isasymA}{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}io\ opt\ {\isasymRightarrow}\ io\ opt{\isachardoublequoteclose}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymA}\ {\isasymphi}\ {\isasymequiv}\ case\ {\isasymphi}\ of\ \isanewline
\ \ \ \ F{\isacharparenleft}{\isasympsi}{\isacharparenright}\ {\isasymRightarrow}\ F{\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasympsi}\ cw{\isacharparenright}\ {\isacharbar}\ P{\isacharparenleft}{\isasympsi}{\isacharparenright}\ {\isasymRightarrow}\ P{\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasympsi}\ cw{\isacharparenright}\ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ ERR{\isacharparenleft}dio{\isacharparenright}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
The Principia Metaphysica distinguishes between encoding and exemplifying, ... say more ...

 Encoding \isa{{\isasymkappa}\isactrlsub {\isadigit{1}}{\isasymPi}\isactrlsup {\isadigit{1}}} is noted here as \isa{{\isasymkappa}\isactrlsub {\isadigit{1}}{\isasymcdot}{\isasymPi}\isactrlsup {\isadigit{1}}}. 
 Encoding yields formulas and never propositional formulas. In the embedding encoding 
 is identified with predicate application.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ Enc{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}e\ opt{\isasymRightarrow}{\isacharparenleft}e{\isasymRightarrow}io{\isacharparenright}\ opt{\isasymRightarrow}io\ opt{\isachardoublequoteclose}{\isacharparenleft}{\isachardoublequoteopen}{\isasymlbrace}{\isacharunderscore}{\isacharcomma}{\isacharunderscore}{\isasymrbrace}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymlbrace}x{\isacharcomma}{\isasymPhi}{\isasymrbrace}\ {\isasymequiv}\ case\ {\isacharparenleft}x{\isacharcomma}{\isasymPhi}{\isacharparenright}\ of\ \isanewline
\ \ \ \ {\isacharparenleft}T{\isacharparenleft}y{\isacharparenright}{\isacharcomma}T{\isacharparenleft}Q{\isacharparenright}{\isacharparenright}\ {\isasymRightarrow}\ F{\isacharparenleft}{\isasymlambda}w{\isachardot}{\isacharparenleft}Q\ y{\isacharparenright}\ w{\isacharparenright}\ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ ERR{\isacharparenleft}dio{\isacharparenright}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
Exemplifying formulas \isa{{\isasymPi}\isactrlsup {\isadigit{1}}{\isasymkappa}\isactrlsub {\isadigit{1}}} are noted here as \isa{{\isasymPi}\isactrlsup {\isadigit{1}}{\isasymbullet}{\isasymkappa}\isactrlsub {\isadigit{1}}}.  
 Exemplification yields propositional formulas and never formulas. In the embedding exemplification 
 is identified with predicate application, just as encoding.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ Exe{\isadigit{1}}{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}e{\isasymRightarrow}io{\isacharparenright}\ opt{\isasymRightarrow}e\ opt{\isasymRightarrow}io\ opt{\isachardoublequoteclose}{\isacharparenleft}{\isachardoublequoteopen}{\isasymlparr}{\isacharunderscore}{\isacharcomma}{\isacharunderscore}{\isasymrparr}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymlparr}{\isasymPhi}{\isacharcomma}x{\isasymrparr}\ {\isasymequiv}\ case\ {\isacharparenleft}{\isasymPhi}{\isacharcomma}x{\isacharparenright}\ of\ \isanewline
\ \ \ \ {\isacharparenleft}T{\isacharparenleft}Q{\isacharparenright}{\isacharcomma}T{\isacharparenleft}y{\isacharparenright}{\isacharparenright}\ {\isasymRightarrow}\ P{\isacharparenleft}{\isasymlambda}w{\isachardot}{\isacharparenleft}Q\ y{\isacharparenright}\ w{\isacharparenright}\ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ ERR{\isacharparenleft}dio{\isacharparenright}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
The Principia Metaphysica supports $n$-ary exemplification constructions. For pragmatical 
 reasons we consider here the cases only for $n\leq 3$.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ Exe{\isadigit{2}}{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}e{\isasymRightarrow}e{\isasymRightarrow}io{\isacharparenright}\ opt{\isasymRightarrow}e\ opt{\isasymRightarrow}e\ opt{\isasymRightarrow}io\ opt{\isachardoublequoteclose}{\isacharparenleft}{\isachardoublequoteopen}{\isasymlparr}{\isacharunderscore}{\isacharcomma}{\isacharunderscore}{\isacharcomma}{\isacharunderscore}{\isasymrparr}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymlparr}{\isasymPhi}{\isacharcomma}x{\isadigit{1}}{\isacharcomma}x{\isadigit{2}}{\isasymrparr}\ {\isasymequiv}\ case\ {\isacharparenleft}{\isasymPhi}{\isacharcomma}x{\isadigit{1}}{\isacharcomma}x{\isadigit{2}}{\isacharparenright}\ of\ \isanewline
\ \ \ \ {\isacharparenleft}T{\isacharparenleft}Q{\isacharparenright}{\isacharcomma}T{\isacharparenleft}y{\isadigit{1}}{\isacharparenright}{\isacharcomma}T{\isacharparenleft}y{\isadigit{2}}{\isacharparenright}{\isacharparenright}\ {\isasymRightarrow}\ P{\isacharparenleft}{\isasymlambda}w{\isachardot}{\isacharparenleft}Q\ y{\isadigit{1}}\ y{\isadigit{2}}{\isacharparenright}\ w{\isacharparenright}\ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ ERR{\isacharparenleft}dio{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ Exe{\isadigit{3}}{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}e{\isasymRightarrow}e{\isasymRightarrow}e{\isasymRightarrow}io{\isacharparenright}\ opt{\isasymRightarrow}e\ opt{\isasymRightarrow}e\ opt{\isasymRightarrow}e\ opt{\isasymRightarrow}io\ opt{\isachardoublequoteclose}{\isacharparenleft}{\isachardoublequoteopen}{\isasymlparr}{\isacharunderscore}{\isacharcomma}{\isacharunderscore}{\isacharcomma}{\isacharunderscore}{\isacharcomma}{\isacharunderscore}{\isasymrparr}{\isachardoublequoteclose}{\isacharparenright}\ \ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymlparr}{\isasymPhi}{\isacharcomma}x{\isadigit{1}}{\isacharcomma}x{\isadigit{2}}{\isacharcomma}x{\isadigit{3}}{\isasymrparr}\ {\isasymequiv}\ case\ {\isacharparenleft}{\isasymPhi}{\isacharcomma}x{\isadigit{1}}{\isacharcomma}x{\isadigit{2}}{\isacharcomma}x{\isadigit{3}}{\isacharparenright}\ of\ \isanewline
\ \ \ \ {\isacharparenleft}T{\isacharparenleft}Q{\isacharparenright}{\isacharcomma}T{\isacharparenleft}y{\isadigit{1}}{\isacharparenright}{\isacharcomma}T{\isacharparenleft}y{\isadigit{2}}{\isacharparenright}{\isacharcomma}T{\isacharparenleft}y{\isadigit{3}}{\isacharparenright}{\isacharparenright}\ {\isasymRightarrow}\ P{\isacharparenleft}{\isasymlambda}w{\isachardot}{\isacharparenleft}Q\ y{\isadigit{1}}\ y{\isadigit{2}}\ y{\isadigit{3}}{\isacharparenright}\ w{\isacharparenright}\ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ ERR{\isacharparenleft}dio{\isacharparenright}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
Formations with negation and implication are supported for both, formulas  and propositional
 formulas, and their embeddings are straightforward.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ not{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}io\ opt{\isasymRightarrow}io\ opt{\isachardoublequoteclose}{\isacharparenleft}{\isachardoublequoteopen}\isactrlbold {\isasymnot}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}\isactrlbold {\isasymnot}\ {\isasymphi}\ {\isasymequiv}\ case\ {\isasymphi}\ of\ \isanewline
\ \ \ \ F{\isacharparenleft}{\isasympsi}{\isacharparenright}\ {\isasymRightarrow}\ F{\isacharparenleft}{\isasymlambda}w{\isachardot}{\isasymnot}{\isacharparenleft}{\isasympsi}\ w{\isacharparenright}{\isacharparenright}\ {\isacharbar}\ P{\isacharparenleft}{\isasympsi}{\isacharparenright}\ {\isasymRightarrow}\ P{\isacharparenleft}{\isasymlambda}w{\isachardot}{\isasymnot}{\isacharparenleft}{\isasympsi}\ w{\isacharparenright}{\isacharparenright}\ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ ERR{\isacharparenleft}dio{\isacharparenright}{\isachardoublequoteclose}\ \ \isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ implies{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}io\ opt{\isasymRightarrow}io\ opt{\isasymRightarrow}io\ opt{\isachardoublequoteclose}{\isacharparenleft}\isakeyword{infixl}\ {\isachardoublequoteopen}\isactrlbold {\isasymrightarrow}{\isachardoublequoteclose}\ {\isadigit{5}}{\isadigit{1}}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymphi}\ \isactrlbold {\isasymrightarrow}\ {\isasympsi}\ {\isasymequiv}\ case\ {\isacharparenleft}{\isasymphi}{\isacharcomma}{\isasympsi}{\isacharparenright}\ of\ \isanewline
\ \ \ \ {\isacharparenleft}F{\isacharparenleft}{\isasymalpha}{\isacharparenright}{\isacharcomma}F{\isacharparenleft}{\isasymbeta}{\isacharparenright}{\isacharparenright}\ {\isasymRightarrow}\ F{\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymalpha}\ w\ {\isasymlongrightarrow}\ {\isasymbeta}\ w{\isacharparenright}\ {\isacharbar}\ {\isacharparenleft}P{\isacharparenleft}{\isasymalpha}{\isacharparenright}{\isacharcomma}P{\isacharparenleft}{\isasymbeta}{\isacharparenright}{\isacharparenright}\ {\isasymRightarrow}\ P{\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymalpha}\ w\ {\isasymlongrightarrow}\ {\isasymbeta}\ w{\isacharparenright}\ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ ERR{\isacharparenleft}dio{\isacharparenright}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
Also universal quantification \isa{\isactrlbold {\isasymforall}{\isacharparenleft}{\isasymlambda}x{\isachardot}{\isasymphi}{\isacharparenright}} (first-order and higher-order) is supported 
 for formulas  and propositional formulas. Following previous work the embedding maps 
 \isa{\isactrlbold {\isasymforall}{\isacharparenleft}{\isasymlambda}x{\isachardot}{\isasymphi}{\isacharparenright}} to \isa{{\isacharparenleft}{\isasymlambda}w{\isachardot}\isactrlbold {\isasymforall}x{\isachardot}{\isasymphi}w{\isacharparenright}}. Note that \isa{\isactrlbold {\isasymforall}} is introduced as logical connective
 based on the existing \isa{{\isasymlambda}}-binder. To improve presentation in the remainder we additional
 introduce binder notation \isa{\isactrlbold {\isasymforall}x{\isachardot}{\isasymphi}} as syntactic sugar for \isa{\isactrlbold {\isasymforall}{\isacharparenleft}{\isasymlambda}x{\isachardot}{\isasymphi}{\isacharparenright}}.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ forall{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}{\isacharprime}a{\isasymRightarrow}io\ opt{\isacharparenright}{\isasymRightarrow}io\ opt{\isachardoublequoteclose}{\isacharparenleft}{\isachardoublequoteopen}\isactrlbold {\isasymforall}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}\isactrlbold {\isasymforall}\ {\isasymPhi}\ {\isasymequiv}\ case\ {\isacharparenleft}{\isasymPhi}\ da{\isacharparenright}\ of\isanewline
\ \ \ \ F{\isacharparenleft}{\isasymphi}{\isacharparenright}\ {\isasymRightarrow}\ F{\isacharparenleft}{\isasymlambda}w{\isachardot}{\isasymforall}x{\isachardot}\ case\ {\isacharparenleft}{\isasymPhi}\ x{\isacharparenright}\ of\ F{\isacharparenleft}{\isasympsi}{\isacharparenright}\ {\isasymRightarrow}\ {\isasympsi}\ w{\isacharparenright}\ {\isacharbar}\ P{\isacharparenleft}{\isasymphi}{\isacharparenright}\ {\isasymRightarrow}\ P{\isacharparenleft}{\isasymlambda}w{\isachardot}{\isasymforall}x{\isachardot}\ case\ {\isacharparenleft}{\isasymPhi}\ x{\isacharparenright}\ of\ P{\isacharparenleft}{\isasympsi}{\isacharparenright}\ {\isasymRightarrow}\ {\isasympsi}\ w{\isacharparenright}\ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ ERR{\isacharparenleft}dio{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ forallBinder{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}{\isacharprime}a{\isasymRightarrow}io\ opt{\isacharparenright}{\isasymRightarrow}io\ opt{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{binder}\ {\isachardoublequoteopen}\isactrlbold {\isasymforall}{\isachardoublequoteclose}\ {\isacharbrackleft}{\isadigit{9}}{\isacharbrackright}\ {\isadigit{9}}{\isacharparenright}\ \ \isakeyword{where}\ {\isachardoublequoteopen}\isactrlbold {\isasymforall}\ x{\isachardot}\ {\isasymphi}\ x\ {\isasymequiv}\ \isactrlbold {\isasymforall}\ {\isasymphi}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\isadelimproof
%
\endisadelimproof
%
\isatagproof
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
The modal \isa{\isactrlbold {\isasymbox}} operator is introduced here for logic S5. Since in an equivalence class
 of possible worlds each world is reachable from any other world, the guarding accessibility clause
 in the usual definition of the \isa{\isactrlbold {\isasymbox}} operator can be omitted. This is convenient and should also
 ease theorem proving. In Section \ref{sec:S5} we will actually demonstrate that the expected S5 properties
 are validated by our modeling of \isa{\isactrlbold {\isasymbox}}.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ box{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}io\ opt{\isasymRightarrow}io\ opt{\isachardoublequoteclose}{\isacharparenleft}{\isachardoublequoteopen}\isactrlbold {\isasymbox}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}\isactrlbold {\isasymbox}{\isasymphi}\ {\isasymequiv}\ case\ {\isasymphi}\ of\ \isanewline
\ \ \ \ F{\isacharparenleft}{\isasympsi}{\isacharparenright}\ {\isasymRightarrow}\ F{\isacharparenleft}{\isasymlambda}w{\isachardot}{\isasymforall}v{\isachardot}\ {\isasympsi}\ v{\isacharparenright}\ {\isacharbar}\ P{\isacharparenleft}{\isasympsi}{\isacharparenright}\ {\isasymRightarrow}\ P{\isacharparenleft}{\isasymlambda}w{\isachardot}{\isasymforall}v{\isachardot}\ {\isasympsi}\ v{\isacharparenright}\ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ ERR{\isacharparenleft}dio{\isacharparenright}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
n-ary lambda abstraction \isa{{\isasymlambda}\isactrlsup {\isadigit{0}}{\isacharcomma}{\isasymlambda}\isactrlsup {\isadigit{1}}{\isacharcomma}{\isasymlambda}\isactrlsup {\isadigit{2}}{\isacharcomma}{\isasymlambda}\isactrlsup {\isadigit{3}}{\isacharcomma}{\isachardot}{\isachardot}{\isachardot}}, for $n\geq 0$, is supported in the Principia
 Metaphysica only for propositional formulas. ... say more about \isa{{\isasymlambda}\isactrlsup {\isadigit{0}}} ... Their embedding is 
 straightforward: \isa{{\isasymlambda}\isactrlsup {\isadigit{0}}} is mapped to identity and \isa{{\isasymlambda}\isactrlsup {\isadigit{1}}{\isacharcomma}{\isasymlambda}\isactrlsup {\isadigit{2}}{\isacharcomma}{\isasymlambda}\isactrlsup {\isadigit{3}}{\isacharcomma}{\isachardot}{\isachardot}{\isachardot}} are mapped to n-ary
 lambda abstractions, that is, \isa{{\isasymlambda}\isactrlsup {\isadigit{2}}{\isacharparenleft}{\isasymlambda}xy{\isachardot}{\isasymphi}{\isacharparenright}} is e.g. mapped to \isa{{\isacharparenleft}{\isasymlambda}xy{\isachardot}{\isasymphi}{\isacharparenright}}.
 For pragmatical reasons we restrict ourselves here to the cases where $n\leq 0$.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ lam{\isadigit{0}}{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}io\ opt{\isasymRightarrow}io\ opt{\isachardoublequoteclose}{\isacharparenleft}{\isachardoublequoteopen}\isactrlbold {\isasymlambda}\isactrlsup {\isadigit{0}}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}\isactrlbold {\isasymlambda}\isactrlsup {\isadigit{0}}{\isasymphi}\ {\isasymequiv}\ case\ {\isasymphi}\ of\ \isanewline
\ \ \ \ P{\isacharparenleft}{\isasympsi}{\isacharparenright}\ {\isasymRightarrow}\ P{\isacharparenleft}{\isasympsi}{\isacharparenright}\ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ ERR\ dio{\isachardoublequoteclose}\ \ \isanewline
\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ lam{\isadigit{1}}{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}e{\isasymRightarrow}io\ opt{\isacharparenright}{\isasymRightarrow}{\isacharparenleft}e{\isasymRightarrow}io{\isacharparenright}\ opt{\isachardoublequoteclose}{\isacharparenleft}{\isachardoublequoteopen}\isactrlbold {\isasymlambda}\isactrlsup {\isadigit{1}}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isacharparenleft}\isactrlbold {\isasymlambda}\isactrlsup {\isadigit{1}}{\isasymPhi}{\isacharparenright}\ {\isasymequiv}\ case\ {\isacharparenleft}{\isasymPhi}\ de{\isacharparenright}\ of\isanewline
\ \ \ \ P{\isacharparenleft}{\isasymphi}{\isacharparenright}\ {\isasymRightarrow}\ T{\isacharparenleft}{\isasymlambda}x{\isachardot}\ case\ {\isacharparenleft}{\isasymPhi}\ x{\isacharparenright}\ of\ P{\isacharparenleft}{\isasymphi}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymphi}{\isacharparenright}\ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ ERR{\isacharparenleft}{\isasymlambda}x{\isachardot}\ dio{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ lam{\isadigit{1}}Binder{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}e{\isasymRightarrow}io\ opt{\isacharparenright}{\isasymRightarrow}{\isacharparenleft}e{\isasymRightarrow}io{\isacharparenright}\ opt{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{binder}\ {\isachardoublequoteopen}\isactrlbold {\isasymlambda}\isactrlsup {\isadigit{1}}{\isachardoublequoteclose}\ {\isacharbrackleft}{\isadigit{9}}{\isacharbrackright}\ {\isadigit{9}}{\isacharparenright}\ \ \isakeyword{where}\ {\isachardoublequoteopen}\isactrlbold {\isasymlambda}\isactrlsup {\isadigit{1}}\ x{\isachardot}\ {\isasymphi}\ x\ {\isasymequiv}\ \isactrlbold {\isasymlambda}\isactrlsup {\isadigit{1}}\ {\isasymphi}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ lam{\isadigit{2}}{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}e{\isasymRightarrow}e{\isasymRightarrow}io\ opt{\isacharparenright}{\isasymRightarrow}{\isacharparenleft}e{\isasymRightarrow}e{\isasymRightarrow}io{\isacharparenright}\ opt{\isachardoublequoteclose}{\isacharparenleft}{\isachardoublequoteopen}\isactrlbold {\isasymlambda}\isactrlsup {\isadigit{2}}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}\isactrlbold {\isasymlambda}\isactrlsup {\isadigit{2}}{\isasymPhi}\ {\isasymequiv}\ case\ {\isacharparenleft}{\isasymPhi}\ de\ de{\isacharparenright}\ of\isanewline
\ \ \ \ P{\isacharparenleft}{\isasymphi}{\isacharparenright}\ {\isasymRightarrow}\ T{\isacharparenleft}{\isasymlambda}x\ y{\isachardot}\ case\ {\isacharparenleft}{\isasymPhi}\ x\ y{\isacharparenright}\ of\ P{\isacharparenleft}{\isasymphi}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymphi}{\isacharparenright}\ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ ERR{\isacharparenleft}{\isasymlambda}x\ y{\isachardot}\ dio{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ lam{\isadigit{3}}{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}e{\isasymRightarrow}e{\isasymRightarrow}e{\isasymRightarrow}io\ opt{\isacharparenright}{\isasymRightarrow}{\isacharparenleft}e{\isasymRightarrow}e{\isasymRightarrow}e{\isasymRightarrow}io{\isacharparenright}\ opt{\isachardoublequoteclose}{\isacharparenleft}{\isachardoublequoteopen}\isactrlbold {\isasymlambda}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}\isactrlbold {\isasymlambda}{\isasymPhi}\ {\isasymequiv}\ case\ {\isacharparenleft}{\isasymPhi}\ de\ de\ de{\isacharparenright}\ of\isanewline
\ \ \ \ P{\isacharparenleft}{\isasymphi}{\isacharparenright}\ {\isasymRightarrow}\ T{\isacharparenleft}{\isasymlambda}x\ y\ z{\isachardot}\ case\ {\isacharparenleft}{\isasymPhi}\ x\ y\ z{\isacharparenright}\ of\ P{\isacharparenleft}{\isasymphi}{\isacharparenright}\ {\isasymRightarrow}\ {\isasymphi}{\isacharparenright}\ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ ERR{\isacharparenleft}{\isasymlambda}x\ y\ z{\isachardot}\ dio{\isacharparenright}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
The Principia Metaphysica supports rigid definite descriptions. Our definition maps
 \isa{{\isasymiota}{\isacharparenleft}{\isasymlambda}x{\isachardot}{\isasymphi}{\isacharparenright}} to \isa{{\isacharparenleft}THE\ x{\isachardot}\ {\isasymphi}\ cw{\isacharparenright}}, that is Isabelle's inbuilt description operator THE 
 is utilized and evaluation is rigidly carried out with respect to the current world \isa{cw}.
 Moreover, application of that is allowed only if the body of \isa{{\isasymPhi}}, computed by clause 
 \isa{{\isacharparenleft}{\isasymPhi}\ de{\isacharparenright}}, is a propositional formula. In this case a term is returned and otherwise error 
 reported (for dummy entity \isa{de}).%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ that{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}e{\isasymRightarrow}io\ opt{\isacharparenright}{\isasymRightarrow}e\ opt{\isachardoublequoteclose}{\isacharparenleft}{\isachardoublequoteopen}{\isasymiota}{\isachardoublequoteclose}{\isacharparenright}\ \ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymiota}\ {\isasymPhi}\ {\isasymequiv}\ case\ {\isacharparenleft}{\isasymPhi}\ de{\isacharparenright}\ of\isanewline
\ \ \ \ P{\isacharparenleft}{\isasymphi}{\isacharparenright}\ {\isasymRightarrow}\ T{\isacharparenleft}THE\ x{\isachardot}\ case\ {\isacharparenleft}{\isasymPhi}\ x{\isacharparenright}\ of\ P\ {\isasympsi}\ {\isasymRightarrow}\ {\isasympsi}\ cw{\isacharparenright}\ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ ERR{\isacharparenleft}de{\isacharparenright}{\isachardoublequoteclose}%
\isamarkupsection{Further Logical Connectives%
}
\isamarkuptrue%
%
\begin{isamarkuptext}%
Further logical connectives can be defined as usual. For existential quantification we here
 prefer a native introduction, even though a definition based on \isa{\isactrlbold {\isasymnot}} and \isa{\isactrlbold {\isasymforall}} is also
 possible (but syntactically not more elegant). For pragmatic reasons it eventually makes sense to 
 to prefer native introductions for all connectives.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ conj{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}io\ opt{\isasymRightarrow}io\ opt{\isasymRightarrow}io\ opt{\isachardoublequoteclose}{\isacharparenleft}\isakeyword{infixl}\ {\isachardoublequoteopen}\isactrlbold {\isasymand}{\isachardoublequoteclose}\ {\isadigit{5}}{\isadigit{3}}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymphi}\isactrlbold {\isasymand}{\isasympsi}\ {\isasymequiv}\ \isactrlbold {\isasymnot}{\isacharparenleft}{\isasymphi}\isactrlbold {\isasymrightarrow}\isactrlbold {\isasymnot}{\isasympsi}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ or{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}io\ opt{\isasymRightarrow}io\ opt{\isasymRightarrow}io\ opt{\isachardoublequoteclose}{\isacharparenleft}\isakeyword{infixl}\ {\isachardoublequoteopen}\isactrlbold {\isasymor}{\isachardoublequoteclose}\ {\isadigit{5}}{\isadigit{2}}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymphi}\isactrlbold {\isasymor}{\isasympsi}\ {\isasymequiv}\ \isactrlbold {\isasymnot}{\isasymphi}\isactrlbold {\isasymrightarrow}{\isasympsi}{\isachardoublequoteclose}\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ equivalent{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}io\ opt{\isasymRightarrow}io\ opt{\isasymRightarrow}io\ opt{\isachardoublequoteclose}{\isacharparenleft}\isakeyword{infixl}\ {\isachardoublequoteopen}\isactrlbold {\isasymequiv}{\isachardoublequoteclose}\ {\isadigit{5}}{\isadigit{2}}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isasymphi}\isactrlbold {\isasymequiv}{\isasympsi}\ {\isasymequiv}\ {\isacharparenleft}{\isasymphi}\isactrlbold {\isasymrightarrow}{\isasympsi}{\isacharparenright}\isactrlbold {\isasymand}{\isacharparenleft}{\isasympsi}\isactrlbold {\isasymrightarrow}{\isasymphi}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ diamond{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}io\ opt{\isasymRightarrow}io\ opt{\isachardoublequoteclose}{\isacharparenleft}{\isachardoublequoteopen}\isactrlbold {\isasymdiamond}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}\isactrlbold {\isasymdiamond}{\isasymphi}\ {\isasymequiv}\ \isactrlbold {\isasymnot}{\isacharparenleft}\isactrlbold {\isasymbox}{\isacharparenleft}\isactrlbold {\isasymnot}\ {\isasymphi}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ exists{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}{\isacharprime}a{\isasymRightarrow}io\ opt{\isacharparenright}{\isasymRightarrow}io\ opt{\isachardoublequoteclose}{\isacharparenleft}{\isachardoublequoteopen}\isactrlbold {\isasymexists}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}\isactrlbold {\isasymexists}{\isasymPhi}\ {\isasymequiv}\ case\ {\isacharparenleft}{\isasymPhi}\ da{\isacharparenright}\ of\isanewline
\ \ \ \ P\ {\isasymphi}\ {\isasymRightarrow}\ P{\isacharparenleft}{\isasymlambda}w{\isachardot}{\isasymexists}x{\isachardot}\ case\ {\isacharparenleft}{\isasymPhi}\ x{\isacharparenright}\ of\ P\ {\isasympsi}\ {\isasymRightarrow}\ {\isasympsi}\ w{\isacharparenright}\ {\isacharbar}\ F\ {\isasymphi}\ {\isasymRightarrow}\ F{\isacharparenleft}{\isasymlambda}w{\isachardot}\ {\isasymexists}x{\isachardot}\ case\ {\isacharparenleft}{\isasymPhi}\ x{\isacharparenright}\ of\ F\ {\isasympsi}\ {\isasymRightarrow}\ {\isasympsi}\ w{\isacharparenright}\ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ ERR\ dio{\isachardoublequoteclose}\ \isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ existsBinder{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}{\isacharprime}a{\isasymRightarrow}io\ opt{\isacharparenright}{\isasymRightarrow}io\ opt{\isachardoublequoteclose}\ {\isacharparenleft}\isakeyword{binder}\ {\isachardoublequoteopen}\isactrlbold {\isasymexists}{\isachardoublequoteclose}\ {\isacharbrackleft}{\isadigit{9}}{\isacharbrackright}\ {\isadigit{9}}{\isacharparenright}\ \ \isakeyword{where}\ {\isachardoublequoteopen}\isactrlbold {\isasymexists}\ x{\isachardot}\ {\isasymphi}\ x\ {\isasymequiv}\ \isactrlbold {\isasymexists}\ {\isasymphi}{\isachardoublequoteclose}%
\isamarkupsection{Meta-Logic%
}
\isamarkuptrue%
%
\begin{isamarkuptext}%
Our approach to rigorously distinguish between proper and improper language constructions 
 and to explicitly maintain respective information is continued also at meta-level. For this 
 we introduce three truth values \isa{tt},
 \isa{ff} and \isa{err}, representing truth, falsity and error. These values can also 
 be noted as \isa{{\isasymtop}}, \isa{{\isasymbottom}} and \isa{{\isacharasterisk}}. We could, of course, also introduce  
 respective logical connectives for the meta-level, but in our applications so far (see below)
 this was not yet relevant.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{datatype}\isamarkupfalse%
\ mf\ {\isacharequal}\ tt\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymtop}{\isachardoublequoteclose}{\isacharparenright}\ {\isacharbar}\ ff\ {\isacharparenleft}{\isachardoublequoteopen}{\isasymbottom}{\isachardoublequoteclose}{\isacharparenright}\ {\isacharbar}\ err\ {\isacharparenleft}{\isachardoublequoteopen}{\isacharasterisk}{\isachardoublequoteclose}{\isacharparenright}%
\begin{isamarkuptext}%
Next we define the meta-logical notions of validity, satisfiability, 
 countersatisfiability and invalidity for our embedded modal relational type theory. To support
 concise formula representations in the remainder we introduce the following notations: \isa{{\isacharbrackleft}{\isasymphi}{\isacharbrackright}} 
 (for \isa{{\isasymphi}} is valid), \isa{{\isacharbrackleft}{\isasymphi}{\isacharbrackright}\isactrlsup s\isactrlsup a\isactrlsup t} (\isa{{\isasymphi}} is satisfiability), \isa{{\isacharbrackleft}{\isasymphi}{\isacharbrackright}\isactrlsup c\isactrlsup s\isactrlsup a\isactrlsup t} 
 (\isa{{\isasymphi}} is countersatisfiability) and \isa{{\isacharbrackleft}{\isasymphi}{\isacharbrackright}\isactrlsup i\isactrlsup n\isactrlsup v} (\isa{{\isasymphi}} is invalid).%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ valid\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}io\ opt{\isasymRightarrow}mf{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isacharbrackleft}{\isacharunderscore}{\isacharbrackright}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymphi}{\isacharbrackright}\ {\isasymequiv}\ case\ {\isasymphi}\ of\ \isanewline
\ \ \ \ P{\isacharparenleft}{\isasympsi}{\isacharparenright}\ {\isasymRightarrow}\ if\ {\isasymforall}w{\isachardot}{\isacharparenleft}{\isasympsi}\ w{\isacharparenright}\ {\isasymlongleftrightarrow}\ True\ then\ {\isasymtop}\ else\ {\isasymbottom}\ {\isacharbar}\ F{\isacharparenleft}{\isasympsi}{\isacharparenright}\ {\isasymRightarrow}\ if\ {\isasymforall}w{\isachardot}{\isacharparenleft}{\isasympsi}\ w{\isacharparenright}\ {\isasymlongleftrightarrow}\ True\ then\ {\isasymtop}\ else\ {\isasymbottom}\ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ {\isacharasterisk}{\isachardoublequoteclose}\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ satisfiable\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}io\ opt{\isasymRightarrow}mf{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isacharbrackleft}{\isacharunderscore}{\isacharbrackright}\isactrlsup s\isactrlsup a\isactrlsup t{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymphi}{\isacharbrackright}\isactrlsup s\isactrlsup a\isactrlsup t\ {\isasymequiv}\ case\ {\isasymphi}\ of\ \isanewline
\ \ \ \ P{\isacharparenleft}{\isasympsi}{\isacharparenright}\ {\isasymRightarrow}\ if\ {\isasymexists}w{\isachardot}{\isacharparenleft}{\isasympsi}\ w{\isacharparenright}\ {\isasymlongleftrightarrow}\ True\ then\ {\isasymtop}\ else\ {\isasymbottom}\ {\isacharbar}\ F{\isacharparenleft}{\isasympsi}{\isacharparenright}\ {\isasymRightarrow}\ if\ {\isasymexists}w{\isachardot}{\isacharparenleft}{\isasympsi}\ w{\isacharparenright}\ {\isasymlongleftrightarrow}\ True\ then\ {\isasymtop}\ else\ {\isasymbottom}\ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ {\isacharasterisk}{\isachardoublequoteclose}\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ countersatisfiable\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}io\ opt{\isasymRightarrow}mf{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isacharbrackleft}{\isacharunderscore}{\isacharbrackright}\isactrlsup c\isactrlsup s\isactrlsup a\isactrlsup t{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymphi}{\isacharbrackright}\isactrlsup c\isactrlsup s\isactrlsup a\isactrlsup t\ {\isasymequiv}\ \ case\ {\isasymphi}\ of\ \isanewline
\ \ \ \ P{\isacharparenleft}{\isasympsi}{\isacharparenright}\ {\isasymRightarrow}\ if\ {\isasymexists}w{\isachardot}{\isasymnot}{\isacharparenleft}{\isasympsi}\ w{\isacharparenright}\ {\isasymlongleftrightarrow}\ True\ then\ {\isasymtop}\ else\ {\isasymbottom}\ {\isacharbar}\ F{\isacharparenleft}{\isasympsi}{\isacharparenright}\ {\isasymRightarrow}\ if\ {\isasymexists}w{\isachardot}{\isasymnot}{\isacharparenleft}{\isasympsi}\ w{\isacharparenright}\ {\isasymlongleftrightarrow}\ True\ then\ {\isasymtop}\ else\ {\isasymbottom}\ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ {\isacharasterisk}{\isachardoublequoteclose}\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ invalid\ {\isacharcolon}{\isacharcolon}\ {\isachardoublequoteopen}io\ opt{\isasymRightarrow}mf{\isachardoublequoteclose}\ {\isacharparenleft}{\isachardoublequoteopen}{\isacharbrackleft}{\isacharunderscore}{\isacharbrackright}\isactrlsup i\isactrlsup n\isactrlsup v{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymphi}{\isacharbrackright}\isactrlsup i\isactrlsup n\isactrlsup v\ {\isasymequiv}\ case\ {\isasymphi}\ of\ \isanewline
\ \ \ \ P{\isacharparenleft}{\isasympsi}{\isacharparenright}\ {\isasymRightarrow}\ if\ {\isasymforall}w{\isachardot}{\isasymnot}{\isacharparenleft}{\isasympsi}\ w{\isacharparenright}\ {\isasymlongleftrightarrow}\ True\ then\ {\isasymtop}\ else\ {\isasymbottom}\ {\isacharbar}\ F{\isacharparenleft}{\isasympsi}{\isacharparenright}\ {\isasymRightarrow}\ if\ {\isasymforall}w{\isachardot}{\isasymnot}{\isacharparenleft}{\isasympsi}\ w{\isacharparenright}\ {\isasymlongleftrightarrow}\ True\ then\ {\isasymtop}\ else\ {\isasymbottom}\ {\isacharbar}\ {\isacharunderscore}\ {\isasymRightarrow}\ {\isacharasterisk}{\isachardoublequoteclose}%
\isamarkupsection{Some Basic Tests%
}
\isamarkuptrue%
%
\isamarkupsubsection{Verifying Necessitation%
}
\isamarkuptrue%
%
\begin{isamarkuptext}%
The next two lemmata show that necessitation holds for arbitrary formulas 
 and arbitrary propositional formulas. We present the lemma in both variants.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ necessitationF{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymphi}\isactrlsup F{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}\ {\isasymlongrightarrow}\ {\isacharbrackleft}\isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ necessitationP{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymphi}\isactrlsup P{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}\ {\isasymlongrightarrow}\ {\isacharbrackleft}\isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\isamarkupsubsection{Modal Collapse is Countersatisfiable%
}
\isamarkuptrue%
%
\begin{isamarkuptext}%
The modelfinder Nitpick constructs a finite countermodel to the assertion
  that modal collaps is valid. Nitpicks countermodel consists of four worlds i1, i2, i3 and i4.
  Moreover, it defines  \isa{{\isasymphi}} to hold only in world i3 and it suggests i3 as the actual 
  world in which to evaluate the conjecture formula. This countermodel is clearly not minimal.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ modalCollapseF{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymphi}\isactrlsup F\ \isactrlbold {\isasymrightarrow}\ \isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\ \isacommand{nitpick}\isamarkupfalse%
%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{oops}\isamarkupfalse%
\ %
\isamarkupcmt{Countermodel by Nitpick%
}
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ modalCollapseP{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymphi}\isactrlsup P\ \isactrlbold {\isasymrightarrow}\ \isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\ \isacommand{nitpick}\isamarkupfalse%
%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{oops}\isamarkupfalse%
\ %
\isamarkupcmt{Countermodel by Nitpick%
}
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
We also tried to prove the statement \isa{{\isacharbrackleft}{\isasymphi}\isactrlsup F\ \isactrlbold {\isasymrightarrow}\ \isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharbrackright}\isactrlsup c\isactrlsup s\isactrlsup a\isactrlsup t\ {\isacharequal}\ {\isasymtop}} with automated reasoners.
 This was however not successful ... (check again and check why).%
\end{isamarkuptext}%
\isamarkuptrue%
%
\isamarkupsubsection{Verifying S5 Principles \label{sec:S5}%
}
\isamarkuptrue%
%
\begin{isamarkuptext}%
\isa{{\isasymbox}} could have been modeled by employing an equivalence relation \isa{r} in a 
 guarding clause. This has been done in previous work. In a simpler, alternative approach we omitted 
 this clause here (since all worlds are reachable from any world in an equivalence relation). The 
 following lemmata, which check various conditions for S5,  ensure that we have indeed 
 modeled S5 correctly.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ axiom{\isacharunderscore}T{\isacharunderscore}P{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}\isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ axiom{\isacharunderscore}T{\isacharunderscore}F{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}\isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ axiom{\isacharunderscore}B{\isacharunderscore}P{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymphi}\isactrlsup P\ \isactrlbold {\isasymrightarrow}\ \isactrlbold {\isasymbox}{\isacharparenleft}\isactrlbold {\isasymdiamond}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ axiom{\isacharunderscore}B{\isacharunderscore}F{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymphi}\isactrlsup F\ \isactrlbold {\isasymrightarrow}\ \isactrlbold {\isasymbox}{\isacharparenleft}\isactrlbold {\isasymdiamond}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ axiom{\isacharunderscore}{\isadigit{4}}{\isacharunderscore}P{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}\isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ \isactrlbold {\isasymdiamond}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ axiom{\isacharunderscore}{\isadigit{4}}{\isacharunderscore}F{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}\isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ \isactrlbold {\isasymdiamond}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ axiom{\isacharunderscore}D{\isacharunderscore}P{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}\isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ \isactrlbold {\isasymbox}{\isacharparenleft}\isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ axiom{\isacharunderscore}D{\isacharunderscore}F{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}\isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ \isactrlbold {\isasymbox}{\isacharparenleft}\isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ axiom{\isacharunderscore}{\isadigit{5}}{\isacharunderscore}P{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}\isactrlbold {\isasymdiamond}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ \isactrlbold {\isasymbox}{\isacharparenleft}\isactrlbold {\isasymdiamond}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ axiom{\isacharunderscore}{\isadigit{5}}{\isacharunderscore}F{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}\isactrlbold {\isasymdiamond}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ \isactrlbold {\isasymbox}{\isacharparenleft}\isactrlbold {\isasymdiamond}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ test{\isacharunderscore}A{\isacharunderscore}P{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}\isactrlbold {\isasymbox}{\isacharparenleft}\isactrlbold {\isasymdiamond}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ \isactrlbold {\isasymdiamond}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ test{\isacharunderscore}A{\isacharunderscore}F{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}\isactrlbold {\isasymbox}{\isacharparenleft}\isactrlbold {\isasymdiamond}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ \isactrlbold {\isasymdiamond}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ test{\isacharunderscore}B{\isacharunderscore}P{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}\isactrlbold {\isasymdiamond}{\isacharparenleft}\isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ \isactrlbold {\isasymdiamond}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ test{\isacharunderscore}B{\isacharunderscore}F{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}\isactrlbold {\isasymdiamond}{\isacharparenleft}\isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ \isactrlbold {\isasymdiamond}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ test{\isacharunderscore}C{\isacharunderscore}P{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}\isactrlbold {\isasymbox}{\isacharparenleft}\isactrlbold {\isasymdiamond}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ \isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\ \isacommand{nitpick}\isamarkupfalse%
%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{oops}\isamarkupfalse%
\ %
\isamarkupcmt{Countermodel by Nitpick%
}
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ test{\isacharunderscore}C{\isacharunderscore}F{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}\isactrlbold {\isasymbox}{\isacharparenleft}\isactrlbold {\isasymdiamond}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ \isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\ \isacommand{nitpick}\isamarkupfalse%
%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{oops}\isamarkupfalse%
\ %
\isamarkupcmt{Countermodel by Nitpick%
}
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ test{\isacharunderscore}D{\isacharunderscore}P{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}\isactrlbold {\isasymdiamond}{\isacharparenleft}\isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ \isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ test{\isacharunderscore}D{\isacharunderscore}F{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}\isactrlbold {\isasymdiamond}{\isacharparenleft}\isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ \isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\isamarkupsubsection{Relations between  Meta-Logical Notions%
}
\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymphi}\isactrlsup P{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}\ {\isasymlongleftrightarrow}\ {\isacharbrackleft}{\isasymphi}\isactrlsup P{\isacharbrackright}\isactrlsup c\isactrlsup s\isactrlsup a\isactrlsup t\ {\isacharequal}\ {\isasymbottom}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymphi}\isactrlsup P{\isacharbrackright}\isactrlsup s\isactrlsup a\isactrlsup t\ {\isacharequal}\ {\isasymtop}\ {\isasymlongleftrightarrow}\ {\isacharbrackleft}{\isasymphi}\isactrlsup P{\isacharbrackright}\isactrlsup i\isactrlsup n\isactrlsup v\ {\isacharequal}\ {\isasymbottom}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymphi}\isactrlsup F{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}\ {\isasymlongleftrightarrow}\ {\isacharbrackleft}{\isasymphi}\isactrlsup F{\isacharbrackright}\isactrlsup c\isactrlsup s\isactrlsup a\isactrlsup t\ {\isacharequal}\ {\isasymbottom}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymphi}\isactrlsup F{\isacharbrackright}\isactrlsup s\isactrlsup a\isactrlsup t\ {\isacharequal}\ {\isasymtop}\ {\isasymlongleftrightarrow}\ {\isacharbrackleft}{\isasymphi}\isactrlsup F{\isacharbrackright}\isactrlsup i\isactrlsup n\isactrlsup v\ {\isacharequal}\ {\isasymbottom}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
However, for terms we have%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymphi}\isactrlsup T{\isacharbrackright}\ {\isacharequal}\ {\isacharasterisk}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymphi}\isactrlsup T{\isacharbrackright}\isactrlsup s\isactrlsup a\isactrlsup t\ {\isacharequal}\ {\isacharasterisk}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymphi}\isactrlsup T{\isacharbrackright}\isactrlsup c\isactrlsup s\isactrlsup a\isactrlsup t\ {\isacharequal}\ {\isacharasterisk}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ \ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymphi}\isactrlsup T{\isacharbrackright}\isactrlsup i\isactrlsup n\isactrlsup v\ {\isacharequal}\ {\isacharasterisk}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\isamarkupsubsection{Testing for the Correct Propagation of Syntactical Category Information%
}
\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}X{\isachardot}\ {\isasymlparr}R\isactrlsup T{\isacharcomma}a\isactrlsup T{\isasymrparr}\ {\isacharequal}\ {\isacharparenleft}X\isactrlsup P{\isacharparenright}\ {\isasymand}\ {\isasymnot}{\isacharparenleft}{\isasymexists}X{\isachardot}\ {\isasymlparr}R\isactrlsup T{\isacharcomma}a\isactrlsup T{\isasymrparr}\ {\isacharequal}\ {\isacharparenleft}X\isactrlsup F{\isacharparenright}{\isacharparenright}\ {\isasymand}\ {\isasymnot}{\isacharparenleft}{\isasymexists}X{\isachardot}\ {\isasymlparr}R\isactrlsup T{\isacharcomma}a\isactrlsup T{\isasymrparr}\ {\isacharequal}\ {\isacharparenleft}X\isactrlsup T{\isacharparenright}{\isacharparenright}\ {\isasymand}\ {\isasymnot}{\isacharparenleft}{\isasymexists}X{\isachardot}\ {\isasymlparr}R\isactrlsup T{\isacharcomma}a\isactrlsup T{\isasymrparr}\ {\isacharequal}\ {\isacharparenleft}X\isactrlsup E{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}X{\isachardot}\ {\isasymlbrace}x\isactrlsup T{\isacharcomma}R\isactrlsup T{\isasymrbrace}\ {\isacharequal}\ {\isacharparenleft}X\isactrlsup F{\isacharparenright}\ {\isasymand}\ {\isasymnot}{\isacharparenleft}{\isasymexists}X{\isachardot}\ {\isasymlbrace}x\isactrlsup T{\isacharcomma}R\isactrlsup T{\isasymrbrace}\ {\isacharequal}\ {\isacharparenleft}X\isactrlsup P{\isacharparenright}{\isacharparenright}\ {\isasymand}\ {\isasymnot}{\isacharparenleft}{\isasymexists}X{\isachardot}\ {\isasymlbrace}x\isactrlsup T{\isacharcomma}R\isactrlsup T{\isasymrbrace}\ {\isacharequal}\ {\isacharparenleft}X\isactrlsup T{\isacharparenright}{\isacharparenright}\ {\isasymand}\ {\isasymnot}{\isacharparenleft}{\isasymexists}X{\isachardot}\ {\isasymlbrace}x\isactrlsup T{\isacharcomma}R\isactrlsup T{\isasymrbrace}\ {\isacharequal}\ {\isacharparenleft}X\isactrlsup E{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
Most importantly, we have that the following language constructing is evaluated as erroneous at validity
 level.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymlparr}\isactrlbold {\isasymlambda}\isactrlsup {\isadigit{1}}x{\isachardot}\ {\isasymlparr}R\isactrlsup T{\isacharcomma}x\isactrlsup T{\isasymrparr}\ \isactrlbold {\isasymrightarrow}\ {\isasymlbrace}x\isactrlsup T{\isacharcomma}R\isactrlsup T{\isasymrbrace}{\isacharcomma}a\isactrlsup T{\isasymrparr}{\isacharbrackright}\ {\isacharequal}\ {\isacharasterisk}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
This is also confirmed as follows in Isabelle: Isabelle simplifies the following expression
 to \isa{dio\isactrlsup E\ {\isacharequal}\ X} (simply move the curse on \isa{simp} to see this).%
\end{isamarkuptext}%
\isamarkuptrue%
%
\begin{isamarkuptext}%
Todo: ... select, adapt, and explain some of examples below ...%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlparr}\isactrlbold {\isasymlambda}\isactrlsup {\isadigit{1}}x{\isachardot}\ {\isasymlparr}R\isactrlsup T{\isacharcomma}x\isactrlsup T{\isasymrparr}\ \isactrlbold {\isasymrightarrow}\ {\isasymlbrace}x\isactrlsup T{\isacharcomma}R\isactrlsup T{\isasymrbrace}{\isacharcomma}a\isactrlsup T{\isasymrparr}\ {\isacharequal}\ X{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{oops}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymexists}X{\isachardot}\ {\isasymlparr}\isactrlbold {\isasymlambda}\isactrlsup {\isadigit{1}}x{\isachardot}\ {\isasymlparr}R\isactrlsup T{\isacharcomma}x\isactrlsup T{\isasymrparr}\ \isactrlbold {\isasymrightarrow}\ {\isasymlparr}R\isactrlsup T{\isacharcomma}x\isactrlsup T{\isasymrparr}{\isacharcomma}a\isactrlsup T{\isasymrparr}\ {\isacharequal}\ {\isacharparenleft}X\isactrlsup P{\isacharparenright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
\ %
\isamarkupcmt{Solution: \isa{{\isacharparenleft}{\isasymlambda}w{\isachardot}\ True{\isacharparenright}\ {\isacharequal}\ X}%
}
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymlparr}\isactrlbold {\isasymlambda}\isactrlsup {\isadigit{1}}x{\isachardot}\ {\isasymlparr}R\isactrlsup T{\isacharcomma}x\isactrlsup T{\isasymrparr}\ \isactrlbold {\isasymrightarrow}\ {\isasymlparr}R\isactrlsup T{\isacharcomma}x\isactrlsup T{\isasymrparr}{\isacharcomma}a\isactrlsup T{\isasymrparr}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isasymlparr}\isactrlbold {\isasymlambda}\isactrlsup {\isadigit{1}}x{\isachardot}\ {\isasymlparr}R\isactrlsup T{\isacharcomma}x\isactrlsup T{\isasymrparr}\ \isactrlbold {\isasymrightarrow}\ {\isasymlparr}R\isactrlsup T{\isacharcomma}x\isactrlsup T{\isasymrparr}{\isacharcomma}a\isactrlsup T{\isasymrparr}\ {\isacharequal}\ {\isacharparenleft}X\isactrlsup P{\isacharparenright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{oops}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}\isactrlbold {\isasymlambda}\isactrlsup {\isadigit{1}}x{\isachardot}\ {\isasymlparr}R\isactrlsup T{\isacharcomma}x\isactrlsup T{\isasymrparr}\ \isactrlbold {\isasymrightarrow}\ {\isasymlbrace}x\isactrlsup T{\isacharcomma}R\isactrlsup T{\isasymrbrace}{\isacharparenright}\ {\isacharequal}\ X{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{oops}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharparenleft}\isactrlbold {\isasymforall}x{\isachardot}\ {\isasymlparr}R\isactrlsup T{\isacharcomma}x\isactrlsup T{\isasymrparr}\ \isactrlbold {\isasymrightarrow}\ {\isasymlparr}R\isactrlsup T{\isacharcomma}x\isactrlsup T{\isasymrparr}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharparenleft}\isactrlbold {\isasymforall}R{\isachardot}\ \isactrlbold {\isasymforall}{\isacharparenleft}{\isasymlambda}x{\isachardot}\ {\isasymlparr}R\isactrlsup T{\isacharcomma}x\isactrlsup T{\isasymrparr}\ \isactrlbold {\isasymrightarrow}\ {\isasymlparr}R\isactrlsup T{\isacharcomma}x\isactrlsup T{\isasymrparr}{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}\isactrlbold {\isasymforall}x{\isachardot}\ {\isasymlparr}R\isactrlsup T{\isacharcomma}x\isactrlsup T{\isasymrparr}\ \isactrlbold {\isasymrightarrow}\ {\isasymlparr}R\isactrlsup T{\isacharcomma}x\isactrlsup T{\isasymrparr}{\isacharparenright}\ {\isacharequal}\ X{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{oops}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharparenleft}\isactrlbold {\isasymforall}x{\isachardot}\ {\isasymlbrace}x\isactrlsup T{\isacharcomma}R\isactrlsup T{\isasymrbrace}\ \isactrlbold {\isasymrightarrow}\ {\isasymlbrace}x\isactrlsup T{\isacharcomma}R\isactrlsup T{\isasymrbrace}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}\isactrlbold {\isasymforall}x{\isachardot}\ {\isasymlbrace}x\isactrlsup T{\isacharcomma}R\isactrlsup T{\isasymrbrace}\ \isactrlbold {\isasymrightarrow}\ {\isasymlbrace}x\isactrlsup T{\isacharcomma}R\isactrlsup T{\isasymrbrace}{\isacharparenright}\ {\isacharequal}\ X{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{oops}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharparenleft}\isactrlbold {\isasymforall}x{\isachardot}\ {\isasymlparr}R\isactrlsup T{\isacharcomma}x\isactrlsup T{\isasymrparr}\ \isactrlbold {\isasymrightarrow}\ {\isasymlbrace}x\isactrlsup T{\isacharcomma}R\isactrlsup T{\isasymrbrace}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isacharasterisk}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharparenleft}\isactrlbold {\isasymforall}x{\isachardot}\ {\isasymlparr}R\isactrlsup T{\isacharcomma}x\isactrlsup T{\isasymrparr}\ \isactrlbold {\isasymrightarrow}\ {\isasymlbrace}x\isactrlsup T{\isacharcomma}R\isactrlsup T{\isasymrbrace}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ X{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{oops}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}\isactrlbold {\isasymforall}x{\isachardot}\ {\isasymlparr}R\isactrlsup T{\isacharcomma}x\isactrlsup T{\isasymrparr}\ \isactrlbold {\isasymrightarrow}\ {\isasymlbrace}x\isactrlsup T{\isacharcomma}R\isactrlsup T{\isasymrbrace}{\isacharparenright}\ {\isacharequal}\ X{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{oops}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharparenleft}\isactrlbold {\isasymforall}R{\isachardot}\ {\isasymlparr}R\isactrlsup T{\isacharcomma}x\isactrlsup T{\isasymrparr}\ \isactrlbold {\isasymrightarrow}\ {\isasymlbrace}x\isactrlsup T{\isacharcomma}R\isactrlsup T{\isasymrbrace}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isacharasterisk}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}\isactrlbold {\isasymforall}R{\isachardot}\ {\isasymlparr}R\isactrlsup T{\isacharcomma}x\isactrlsup T{\isasymrparr}\ \isactrlbold {\isasymrightarrow}\ {\isasymlbrace}x\isactrlsup T{\isacharcomma}R\isactrlsup T{\isasymrbrace}{\isacharparenright}\ {\isacharequal}\ X{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{oops}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\isamarkupsubsection{Are Priorities Defined Correctly?%
}
\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymand}\ {\isacharparenleft}{\isasympsi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isasymchi}\isactrlsup P{\isacharparenright}\ {\isasymequiv}\ {\isacharparenleft}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymand}\ {\isacharparenleft}{\isasympsi}\isactrlsup P{\isacharparenright}{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isasymchi}\isactrlsup P{\isacharparenright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymand}\ {\isacharparenleft}{\isasympsi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isasymchi}\isactrlsup P{\isacharparenright}\ {\isasymequiv}\ {\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymand}\ {\isacharparenleft}{\isacharparenleft}{\isasympsi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isasymchi}\isactrlsup P{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\ \isacommand{nitpick}\isamarkupfalse%
%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{oops}\isamarkupfalse%
\ %
\isamarkupcmt{Countermodel by Nitpick%
}
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymand}\ {\isacharparenleft}{\isasympsi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymequiv}\ {\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymand}\ {\isacharparenleft}{\isasympsi}\isactrlsup P{\isacharparenright}{\isacharparenright}\ {\isasymequiv}\ {\isacharparenleft}{\isacharparenleft}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymand}\ {\isacharparenleft}{\isasympsi}\isactrlsup P{\isacharparenright}{\isacharparenright}\ \isactrlbold {\isasymequiv}\ {\isacharparenleft}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymand}\ {\isacharparenleft}{\isasympsi}\isactrlsup P{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharparenleft}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymand}\ {\isacharparenleft}{\isasympsi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymequiv}\ {\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymand}\ {\isacharparenleft}{\isasympsi}\isactrlsup P{\isacharparenright}{\isacharparenright}\ {\isasymequiv}\ {\isacharparenleft}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymand}\ {\isacharparenleft}{\isacharparenleft}{\isasympsi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymequiv}\ {\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharparenright}\ \isactrlbold {\isasymand}\ {\isacharparenleft}{\isasympsi}\isactrlsup P{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\ \isacommand{nitpick}\isamarkupfalse%
%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{oops}\isamarkupfalse%
\ %
\isamarkupcmt{Countermodel by Nitpick%
}
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\isamarkupsection{E!, O!, A! and =E%
}
\isamarkuptrue%
%
\begin{isamarkuptext}%
We introduce the distinguished 1-place relation constant: E (read: ‘being concrete’ or ‘concreteness’)%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{consts}\isamarkupfalse%
\ E{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}e{\isasymRightarrow}io{\isacharparenright}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
Being ordinary is defined as being possibly concrete.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ ordinaryObject{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}e{\isasymRightarrow}io{\isacharparenright}\ opt{\isachardoublequoteclose}{\isacharparenleft}{\isachardoublequoteopen}O{\isacharbang}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}O{\isacharbang}\ {\isasymequiv}\ \isactrlbold {\isasymlambda}\isactrlsup {\isadigit{1}}x{\isachardot}\ \isactrlbold {\isasymdiamond}{\isasymlparr}E\isactrlsup T{\isacharcomma}x\isactrlsup T{\isasymrparr}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
Being abstract is is defined as not possibly being concrete.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ abstractObject{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}e{\isasymRightarrow}io{\isacharparenright}\ opt{\isachardoublequoteclose}{\isacharparenleft}{\isachardoublequoteopen}A{\isacharbang}{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}A{\isacharbang}\ {\isasymequiv}\ \isactrlbold {\isasymlambda}\isactrlsup {\isadigit{1}}x{\isachardot}\ \isactrlbold {\isasymnot}{\isacharparenleft}\isactrlbold {\isasymdiamond}{\isasymlparr}E\isactrlsup T{\isacharcomma}x\isactrlsup T{\isasymrparr}{\isacharparenright}{\isachardoublequoteclose}%
\begin{isamarkuptext}%
Identity \isa{{\isacharequal}\isactrlsub e} is defined as .... Subsequenty, the infix operator \isa{{\isacharequal}\isactrlsub E} is 
 introduced.%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{abbreviation}\isamarkupfalse%
\ identity{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}{\isacharparenleft}e{\isasymRightarrow}e{\isasymRightarrow}io{\isacharparenright}\ opt{\isachardoublequoteclose}{\isacharparenleft}{\isachardoublequoteopen}\isactrlbold {\isacharequal}\isactrlsub e{\isachardoublequoteclose}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}\isactrlbold {\isacharequal}\isactrlsub e\ {\isasymequiv}\ \isanewline
\ \ \isactrlbold {\isasymlambda}\isactrlsup {\isadigit{2}}{\isacharparenleft}{\isasymlambda}x\ y{\isachardot}\ {\isacharparenleft}{\isasymlparr}O{\isacharbang}{\isacharcomma}x\isactrlsup T{\isasymrparr}\ \isactrlbold {\isasymand}\ {\isasymlparr}O{\isacharbang}{\isacharcomma}y\isactrlsup T{\isasymrparr}{\isacharparenright}\ \isactrlbold {\isasymand}\ \isactrlbold {\isasymbox}{\isacharparenleft}\isactrlbold {\isasymforall}{\isacharparenleft}{\isasymlambda}F{\isachardot}\ {\isasymlparr}F\isactrlsup T{\isacharcomma}x\isactrlsup T{\isasymrparr}\ \isactrlbold {\isasymequiv}\ {\isasymlparr}F\isactrlsup T{\isacharcomma}y\isactrlsup T{\isasymrparr}{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isanewline
\isacommand{abbreviation}\isamarkupfalse%
\ identityE{\isacharcolon}{\isacharcolon}{\isachardoublequoteopen}e\ opt{\isasymRightarrow}e\ opt{\isasymRightarrow}io\ opt{\isachardoublequoteclose}{\isacharparenleft}\isakeyword{infixl}\ {\isachardoublequoteopen}\isactrlbold {\isacharequal}\isactrlsub E{\isachardoublequoteclose}\ {\isadigit{6}}{\isadigit{3}}{\isacharparenright}\ \isakeyword{where}\ {\isachardoublequoteopen}x\ \isactrlbold {\isacharequal}\isactrlsub E\ y\ {\isasymequiv}\ {\isasymlparr}\isactrlbold {\isacharequal}\isactrlsub e{\isacharcomma}x{\isacharcomma}y{\isasymrparr}{\isachardoublequoteclose}%
\isamarkupsection{Further Test Examples%
}
\isamarkuptrue%
%
\begin{isamarkuptext}%
Todo: ... select, adapt, and explain some of examples below%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharparenleft}\isactrlbold {\isasymforall}x{\isachardot}\ \isactrlbold {\isasymexists}{\isacharparenleft}{\isasymlambda}R{\isachardot}\ {\isasymlbrace}x\isactrlsup T{\isacharcomma}R\isactrlsup T{\isasymrbrace}\ \isactrlbold {\isasymrightarrow}\ {\isasymlbrace}x\isactrlsup T{\isacharcomma}R\isactrlsup T{\isasymrbrace}{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharparenleft}\isactrlbold {\isasymexists}x{\isachardot}\ \isactrlbold {\isasymforall}{\isacharparenleft}{\isasymlambda}R{\isachardot}\ {\isasymlbrace}x\isactrlsup T{\isacharcomma}R\isactrlsup T{\isasymrbrace}\ \isactrlbold {\isasymrightarrow}\ {\isasymlbrace}x\isactrlsup T{\isacharcomma}R\isactrlsup T{\isasymrbrace}{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharparenleft}a\isactrlsup T{\isacharparenright}\ \isactrlbold {\isacharequal}\isactrlsub E\ {\isacharparenleft}a\isactrlsup T{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\ \isacommand{nitpick}\isamarkupfalse%
%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{oops}\isamarkupfalse%
\ %
\isamarkupcmt{Countermodel by Nitpick%
}
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymlparr}O{\isacharbang}{\isacharcomma}a\isactrlsup T{\isasymrparr}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}a\isactrlsup T{\isacharparenright}\ \isactrlbold {\isacharequal}\isactrlsub E\ {\isacharparenleft}a\isactrlsup T{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharparenleft}\isactrlbold {\isasymforall}F{\isachardot}\ {\isasymlparr}F\isactrlsup T{\isacharcomma}x\isactrlsup T{\isasymrparr}\ \isactrlbold {\isasymequiv}\ {\isasymlparr}F\isactrlsup T{\isacharcomma}x\isactrlsup T{\isasymrparr}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymlparr}O{\isacharbang}{\isacharcomma}a\isactrlsup T{\isasymrparr}\ \isactrlbold {\isasymrightarrow}\ {\isasymlparr}{\isacharparenleft}\isactrlbold {\isasymlambda}\isactrlsup {\isadigit{1}}x{\isachardot}\ {\isacharparenleft}x\isactrlsup T{\isacharparenright}\ \isactrlbold {\isacharequal}\isactrlsub E\ {\isacharparenleft}a\isactrlsup T{\isacharparenright}{\isacharparenright}{\isacharcomma}a\isactrlsup T{\isasymrparr}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharparenleft}\isactrlbold {\isasymexists}F{\isachardot}\ {\isasymlbrace}a\isactrlsup T{\isacharcomma}F\isactrlsup T{\isasymrbrace}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharparenleft}\isactrlbold {\isasymexists}{\isasymphi}{\isachardot}\ {\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharparenleft}\isactrlbold {\isasymexists}{\isasymphi}{\isachardot}\ {\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{by}\isamarkupfalse%
\ auto%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\isamarkupsection{Axioms%
}
\isamarkuptrue%
%
\isamarkupsubsection{Axioms for Negations and Conditionals%
}
\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ a{\isadigit{2}}{\isadigit{1}}{\isacharunderscore}{\isadigit{1}}{\isacharunderscore}P{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ a{\isadigit{2}}{\isadigit{1}}{\isacharunderscore}{\isadigit{1}}{\isacharunderscore}F{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ a{\isadigit{2}}{\isadigit{1}}{\isacharunderscore}{\isadigit{2}}{\isacharunderscore}P{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharparenleft}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isacharparenleft}{\isasympsi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isasymchi}\isactrlsup P{\isacharparenright}{\isacharparenright}{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isacharparenleft}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isasympsi}\isactrlsup P{\isacharparenright}{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isasymchi}\isactrlsup P{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ a{\isadigit{2}}{\isadigit{1}}{\isacharunderscore}{\isadigit{2}}{\isacharunderscore}F{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharparenleft}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isacharparenleft}{\isasympsi}\isactrlsup F{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isasymchi}\isactrlsup F{\isacharparenright}{\isacharparenright}{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isacharparenleft}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isasympsi}\isactrlsup F{\isacharparenright}{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isasymchi}\isactrlsup F{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ a{\isadigit{2}}{\isadigit{1}}{\isacharunderscore}{\isadigit{3}}{\isacharunderscore}P{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharparenleft}\isactrlbold {\isasymnot}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ \isactrlbold {\isasymnot}{\isacharparenleft}{\isasympsi}\isactrlsup P{\isacharparenright}{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isacharparenleft}\isactrlbold {\isasymnot}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isasympsi}\isactrlsup P{\isacharparenright}{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ a{\isadigit{2}}{\isadigit{1}}{\isacharunderscore}{\isadigit{3}}{\isacharunderscore}F{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharparenleft}\isactrlbold {\isasymnot}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ \isactrlbold {\isasymnot}{\isacharparenleft}{\isasympsi}\isactrlsup F{\isacharparenright}{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isacharparenleft}\isactrlbold {\isasymnot}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isasympsi}\isactrlsup F{\isacharparenright}{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\isamarkupsubsection{Axioms of Identity%
}
\isamarkuptrue%
%
\begin{isamarkuptext}%
todo%
\end{isamarkuptext}%
\isamarkuptrue%
%
\isamarkupsubsection{Axioms of Quantification%
}
\isamarkuptrue%
%
\begin{isamarkuptext}%
todo%
\end{isamarkuptext}%
\isamarkuptrue%
%
\isamarkupsubsection{Axioms of Actuality%
}
\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ a{\isadigit{3}}{\isadigit{1}}{\isacharunderscore}{\isadigit{1}}{\isacharunderscore}P{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymA}{\isacharparenleft}\isactrlbold {\isasymnot}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharparenright}\ \isactrlbold {\isasymequiv}\ \isactrlbold {\isasymnot}{\isacharparenleft}{\isasymA}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ a{\isadigit{3}}{\isadigit{1}}{\isacharunderscore}{\isadigit{1}}{\isacharunderscore}F{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymA}{\isacharparenleft}\isactrlbold {\isasymnot}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharparenright}\ \isactrlbold {\isasymequiv}\ \isactrlbold {\isasymnot}{\isacharparenleft}{\isasymA}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ a{\isadigit{3}}{\isadigit{1}}{\isacharunderscore}{\isadigit{2}}{\isacharunderscore}P{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymA}{\isacharparenleft}{\isasymphi}\isactrlsup P\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isasympsi}\isactrlsup P{\isacharparenright}{\isacharparenright}\ \isactrlbold {\isasymequiv}\ {\isacharparenleft}{\isasymA}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isasymA}{\isacharparenleft}{\isasympsi}\isactrlsup P{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ a{\isadigit{3}}{\isadigit{1}}{\isacharunderscore}{\isadigit{2}}{\isacharunderscore}F{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymA}{\isacharparenleft}{\isasymphi}\isactrlsup F\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isasympsi}\isactrlsup F{\isacharparenright}{\isacharparenright}\ \isactrlbold {\isasymequiv}\ {\isacharparenleft}{\isasymA}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isasymA}{\isacharparenleft}{\isasympsi}\isactrlsup F{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ a{\isadigit{3}}{\isadigit{1}}{\isacharunderscore}{\isadigit{3}}{\isacharunderscore}P{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymA}{\isacharparenleft}\isactrlbold {\isasymforall}x{\isachardot}\ {\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharparenright}\ \isactrlbold {\isasymequiv}\ {\isacharparenleft}\isactrlbold {\isasymforall}x{\isachardot}\ {\isasymA}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ a{\isadigit{3}}{\isadigit{1}}{\isacharunderscore}{\isadigit{3}}{\isacharunderscore}F{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharparenleft}{\isasymA}{\isacharparenleft}\isactrlbold {\isasymforall}x{\isachardot}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharparenright}\ \isactrlbold {\isasymequiv}\ {\isacharparenleft}\isactrlbold {\isasymforall}x{\isachardot}\ {\isasymA}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ a{\isadigit{3}}{\isadigit{1}}{\isacharunderscore}{\isadigit{4}}{\isacharunderscore}P{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymA}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymequiv}\ {\isasymA}{\isacharparenleft}{\isasymA}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ a{\isadigit{3}}{\isadigit{1}}{\isacharunderscore}{\isadigit{4}}{\isacharunderscore}F{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymA}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}\ \isactrlbold {\isasymequiv}\ {\isasymA}{\isacharparenleft}{\isasymA}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\isamarkupsubsection{Axioms of Necessity%
}
\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ a{\isadigit{3}}{\isadigit{2}}{\isacharunderscore}{\isadigit{1}}{\isacharunderscore}P{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharparenleft}\isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup P\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharparenright}{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}\isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ \isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\ \ \ \ \ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ a{\isadigit{3}}{\isadigit{2}}{\isacharunderscore}{\isadigit{1}}{\isacharunderscore}F{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharparenleft}\isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup F\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharparenright}{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}\isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ \isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\ \ \ \ \ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ a{\isadigit{3}}{\isadigit{2}}{\isacharunderscore}{\isadigit{2}}{\isacharunderscore}P{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}\isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ a{\isadigit{3}}{\isadigit{2}}{\isacharunderscore}{\isadigit{2}}{\isacharunderscore}F{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}\isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ {\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isanewline
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ a{\isadigit{3}}{\isadigit{2}}{\isacharunderscore}{\isadigit{3}}{\isacharunderscore}P{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}\isactrlbold {\isasymbox}{\isacharparenleft}\isactrlbold {\isasymdiamond}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ \isactrlbold {\isasymdiamond}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ a{\isadigit{3}}{\isadigit{2}}{\isacharunderscore}{\isadigit{3}}{\isacharunderscore}F{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}\isactrlbold {\isasymbox}{\isacharparenleft}\isactrlbold {\isasymdiamond}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ \isactrlbold {\isasymdiamond}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ a{\isadigit{3}}{\isadigit{2}}{\isacharunderscore}{\isadigit{4}}{\isacharunderscore}P{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharparenleft}\isactrlbold {\isasymforall}x{\isachardot}\ \isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ \isactrlbold {\isasymbox}{\isacharparenleft}{\isacharparenleft}\isactrlbold {\isasymforall}x{\isachardot}\ {\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\ \ \ \ \ \ \ \ \ \isanewline
\isacommand{lemma}\isamarkupfalse%
\ a{\isadigit{3}}{\isadigit{2}}{\isacharunderscore}{\isadigit{4}}{\isacharunderscore}F{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isacharparenleft}\isactrlbold {\isasymforall}x{\isachardot}\ \isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ \isactrlbold {\isasymbox}{\isacharparenleft}{\isacharparenleft}\isactrlbold {\isasymforall}x{\isachardot}\ {\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
The following needs to be an axiom; it does not follow for free: it is possible that there 
are contingently concrete individuals and it is possible that there are not:%
\end{isamarkuptext}%
\isamarkuptrue%
\isacommand{axiomatization}\isamarkupfalse%
\ \isakeyword{where}\isanewline
\ \ a{\isadigit{3}}{\isadigit{2}}{\isacharunderscore}{\isadigit{5}}{\isacharunderscore}P{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}\isactrlbold {\isasymdiamond}{\isacharparenleft}\isactrlbold {\isasymexists}x{\isachardot}\ {\isasymlparr}E\isactrlsup T{\isacharcomma}x\isactrlsup T{\isasymrparr}\ \isactrlbold {\isasymand}\ \isactrlbold {\isasymdiamond}{\isacharparenleft}\isactrlbold {\isasymnot}{\isasymlparr}E\isactrlsup T{\isacharcomma}x\isactrlsup T{\isasymrparr}{\isacharparenright}{\isacharparenright}\ \isactrlbold {\isasymand}\ \isactrlbold {\isasymdiamond}{\isacharparenleft}\isactrlbold {\isasymnot}{\isacharparenleft}\isactrlbold {\isasymexists}x{\isachardot}\ {\isasymlparr}E\isactrlsup T{\isacharcomma}x\isactrlsup T{\isasymrparr}\ \isactrlbold {\isasymand}\ \isactrlbold {\isasymdiamond}{\isacharparenleft}\isactrlbold {\isasymnot}{\isasymlparr}E\isactrlsup T{\isacharcomma}x\isactrlsup T{\isasymrparr}{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isamarkupsubsection{Axioms of Necessity and Actuality%
}
\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ a{\isadigit{3}}{\isadigit{3}}{\isacharunderscore}{\isadigit{1}}{\isacharunderscore}P{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymA}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ \isactrlbold {\isasymbox}{\isacharparenleft}{\isasymA}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ a{\isadigit{3}}{\isadigit{3}}{\isacharunderscore}{\isadigit{1}}{\isacharunderscore}F{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymA}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}\ \isactrlbold {\isasymrightarrow}\ \isactrlbold {\isasymbox}{\isacharparenleft}{\isasymA}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ a{\isadigit{3}}{\isadigit{3}}{\isacharunderscore}{\isadigit{2}}{\isacharunderscore}P{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}\isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}\ \isactrlbold {\isasymequiv}\ {\isacharparenleft}{\isasymA}{\isacharparenleft}\isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup P{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
\isanewline
\isacommand{lemma}\isamarkupfalse%
\ a{\isadigit{3}}{\isadigit{3}}{\isacharunderscore}{\isadigit{2}}{\isacharunderscore}F{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}\isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}\ \isactrlbold {\isasymequiv}\ {\isacharparenleft}{\isasymA}{\isacharparenleft}\isactrlbold {\isasymbox}{\isacharparenleft}{\isasymphi}\isactrlsup F{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isacharbrackright}\ {\isacharequal}\ {\isasymtop}{\isachardoublequoteclose}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{apply}\isamarkupfalse%
\ simp\ \isacommand{done}\isamarkupfalse%
\isanewline
\isanewline
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
\end{isabellebody}%
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "root"
%%% End: