### 2018-09-11 FreeLogic/2018-AFP/AxiomaticCategoryTheory/output/document/root.tex
added

\documentclass[11pt,a4paper]{article}
\usepackage{isabelle,isabellesym}
\usepackage{authblk}
\usepackage{latexsym,amsmath}
\usepackage[margin=2cm]{geometry}

% further packages required for unusual symbols (see also
% isabellesym.sty), use only when needed

\usepackage{amssymb}
  %for \<leadsto>, \<box>, \<diamond>, \<sqsupset>, \<mho>, \<Join>,
  %\<lhd>, \<lesssim>, \<greatersim>, \<lessapprox>, \<greaterapprox>,
  %\<triangleq>, \<yen>, \<lozenge>

%\usepackage{eurosym}
  %for \<euro>

%\usepackage[only,bigsqcap]{stmaryrd}
  %for \<Sqinter>

%\usepackage{eufrak}
  %for \<AA> ... \<ZZ>, \<aa> ... \<zz> (also included in amssymb)

%\usepackage{textcomp}
  %for \<onequarter>, \<onehalf>, \<threequarters>, \<degree>, \<cent>,
  %\<currency>

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

% for uniform font size
%\renewcommand{\isastyle}{\isastyleminor}

\begin{document}

\title{Axiom Systems for Category Theory in Free Logic}
\author[1]{Christoph Benzm\"uller}
\author[2]{Dana S. Scott}
\affil[1]{University of Luxembourg, Luxembourg \& Freie Universit\"at Berlin, Germany}
\affil[2]{Visiting Scholar at University of Califormia, Berkeley,  USA}
\date{\today}                     %% if you don't need date to appear
\setcounter{Maxaffil}{0}
\renewcommand\Affilfont{\itshape\small}

\maketitle

\tableofcontents

% sane default for proof documents
\parindent 0pt\parskip 0.5ex

% generated text of all theories
\input{session}

% optional bibliography
\bibliographystyle{abbrv}
% \bibliography{root}

\begin{thebibliography}{1}

\bibitem{C57}
C.~Benzm{\"u}ller and D.~Scott.
\newblock Automating free logic in {Isabelle/HOL}.
\newblock In G.-M. Greuel, T.~Koch, P.~Paule, and A.~Sommese, editors, {\em
  Mathematical Software -- ICMS 2016, 5th International Congress, Proceedings},
  volume 9725 of {\em LNCS}, pages 43--50, Berlin, Germany, 2016. Springer.

\bibitem{C67}
C.~Benzm{\"u}ller and D.~Scott.
\newblock Some reflections on a computer-aided theory exploration study in
  category theory (extended abstract).
\newblock In T.~C. Hales, C.~Kaliszyk, S.~Schulz, and J.~Urban, editors, {\em
  3rd Conference on Artificial Intelligence and Theorem Proving (AITP 2018),
  Book of Abstracts}, 2018.

\bibitem{R58}
C.~Benzm{\"u}ller and D.~S. Scott.
\newblock Axiomatizing category theory in free logic.
\newblock Technical report, CoRR, 2016.
\newblock \url{http://arxiv.org/abs/1609.01493}.

\bibitem{FreydScedrov90}
P.~J. Freyd and A.~Scedrov.
\newblock {\em Categories, Allegories}.
\newblock North Holland, 1990.

\bibitem{MacLane48}
S.~McLane.
\newblock Groups, categories and duality.
\newblock {\em Proceedings of the National Academy of Sciences},
  34(6):263--267, 1948.

\bibitem{Scott79}
D.~Scott.
\newblock Identity and existence in intuitionistic logic.
\newblock In M.~Fourman, C.~Mulvey, and D.~Scott, editors, {\em Applications of
  Sheaves: Proceedings of the Research Symposium on Applications of Sheaf
  Theory to Logic, Algebra, and Analysis, Durham, July 9--21, 1977}, volume 752
  of {\em Lecture Notes in Mathematics}, pages 660--696. Springer Berlin
  Heidelberg, 1979.

\end{thebibliography}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: